---
output: github_document
---

<!-- badges: start -->
[![R build status](https://github.com/geocompx/geocompkg/workflows/pkgdown/badge.svg)](https://github.com/geocompx/geocompkg/actions)
[![Codecov test coverage](https://codecov.io/gh/geocompx/geocompkg/branch/master/graph/badge.svg)](https://codecov.io/gh/geocompx/geocompkg?branch=master)
[![R build status](https://github.com/geocompx/geocompkg/workflows/R-CMD-check/badge.svg)](https://github.com/geocompx/geocompkg/actions)
[![R-CMD-check](https://github.com/geocompx/geocompkg/workflows/R-CMD-check/badge.svg)](https://github.com/geocompx/geocompkg/actions)
[![R-CMD-check](https://github.com/geocompx/geocompkg/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/geocompx/geocompkg/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geocompkg

This is a 'metapackage' support the book 
[Gecomputation with R](https://r.geocompx.org/).
Installing installing it will give you the dependencies needed to run the code in the book.
The repository also contains up-to-date solutions additional materials to supplement the book.
See [the vignettes folder](https://github.com/geocompx/geocompkg/tree/master/vignettes) for the source code.

## Installation

You can install the packages needed to reproduce the code in Geocompuation with R by installing the [geocompkg](https://github.com/geocompx/geocompkg) package as follows:

```{r, eval=FALSE}
devtools::install_github("geocompx/geocompkg")
```

If you have any issues, check-out [Chapter 2](https://r.geocompx.org/spatial-class.html) of the book.
You may need to install **spDataLarge** from drat:

```{r, eval=FALSE}
install.packages("spDataLarge", repos = "https://nowosad.github.io/drat/",
                 type = "source")
```


```{r, eval=FALSE, echo=FALSE}
# build site:
pkgdown::build_home()

# build vignettes one by one, e.g.
pkgdown::build_article("sea-level-rise")

# pkgdown::build_articles() # for all
# build all articles
articles = list.files(path = "vignettes/", pattern = ".Rmd")
articles = gsub(pattern = ".Rmd", replacement = "", articles)
for(i in articles) {
  pkgdown::build_article(i)
}
```

