---
output: github_document
---

[![Build Status](https://travis-ci.org/geocompr/geocompkg.svg?branch=master)](https://travis-ci.org/geocompr/geocompkg)


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geocompkg

This is a metapackage that will install all packages used in the book 
[Gecomputation with R](https://geocompr.robinlovelace.net/), the source code of which can be found at [github.com/Robinlovelace/geocompr](https://github.com/Robinlovelace/geocompr).
It is used to keep up-to-date solutions additional materials to supplement the book.
See [the vignettes folder](https://github.com/geocompr/geocompkg/tree/master/vignettes) for the source code.

## Installation

You can install the packages needed to reproduce the code in Geocompuation with R by installing the [geocompkg](https://github.com/geocompr/geocompkg) package as follows:

```{r, eval=FALSE}
install.packages("spDataLarge", repos = "https://nowosad.github.io/drat/",
                 type = "source")
devtools::install_github("geocompr/geocompkg")
```


```{r, eval=FALSE, echo=FALSE}
# build all articles
articles = list.files(path = "vignettes/", pattern = ".Rmd")
articles = gsub(pattern = ".Rmd", replacement = "", articles)
for(i in articles) {
  pkgdown::build_article(i)
}
```

