---
title: "Loading and analysing GPS tracks with R"
author: "Robin Lovelace"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Perhaps the most ubiquitous type of geographic information is the continuous stream of data churned-out by GPS devices.
Global Positioning System (GPS) devices are now in everything from watches to cars and, of course, smartphones.
This means that GPS datasets have the ability to track a large proportion of the world's population.
Although there are privacy concerns, when appropriately anonymized and aggregated, GPS datasets have the potential to help tackle the issues raised in Chapter [12](http://geocompr.robinlovelace.net/transport.html) of Gecomputation with R:
to design healthy transport systems in which walking and cycling overtake cars as the main travel modes.

## Loading GPS data

But how to load GPS data?
Because GPS traces are vector datasets, we use **sf** to load them:

```{r}
library(sf)
```


As with any dataset the first stage is to identify the source.
For my own GPS data I use the open source Android app Orux to log my routes which I then transfer onto my computer, into the `Gps` folder in my home directory.
The following code allows me to load the route I took on the 30^th^ August 2018, for example (not run):

```{r, eval=FALSE}
file_gps = "~/Gps/2018-08-30 1515__20180830_1515.gpx"
st_layers(file_gps)
r = st_read(file_gps, layer = "tracks")
plot(r)
mapview::mapview(r)
```

