---
title: "Geocomputation for sea level research"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-slr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Sea level rise (SLR) is a worldwide phenemona with major implications for human civilisation.
Despite the globally interconnected nature of the oceans, the rates and likely impacts of SLR have a strong spatial footprint.
Some countries such as Norway are actually increasing in height and have limited exposure while in some, such as Bangladesh, millions of people are at risk from catastrophic SLR-related flooding.
These imacts are exacerbated by climate change, adding an element of uncertainty and dynamics to the problem.
All these features of SLR make it well-suited to analysis using geocomputation.


```{r}
# aim: demo r-spatial capabilities with slr data
# See https://github.com/geocompr/geostats_18/blob/master/code/geocompr/slr.R

# if(!require(elevatr)) {
#   install.packages("elevatr")
# }
# library(elevatr)
library(raster)
library(sf)
```

```{r}
# load data
if(!dir.exists("data")) {
  download.file("https://github.com/geocompr/geostats_18/releases/download/0.1/data.zip", "data.zip")
  unzip("data.zip")
  file.rename("pres/geocompr/data/", "data")
}
```

```{r, eval=FALSE}
# vector data
plot(spData::world)
poland = dplyr::filter(spData::world, grepl(pattern = "Pola", name_long))
plot(st_geometry(poland))
# testing elevatr (from readme / vignette)
# data(lake)
# elevation_high = get_elev_raster(lake, z = 14, src = "aws")
# 120 MB"
# elevation = get_elev_raster(lake, z = 9, src = "aws")
# elevation_low = get_elev_raster(lake, z = 4, src = "aws")
# pryr::object_size(elevation)
# pryr::object_size(elevation_low)
# plot(elevation_low)
# mapview::mapview(elevation_low)
# cols = rainbow(5)
# plot(elevation, add = T, col = cols)

# getting a map of the szczecin coastline
# option 1 - via osmdata:
# library(osmdata)
# region = osmdata::getbb(place_name = "cardiff uk", format_out = "sf_polygon")
# region = getbb(place_name = "cardiff uk", format_out = "sf_polygon")
# sf::st_write(region, "data/scz-osm.geojson")
region = sf::read_sf("data/scz-osm.geojson")
(m = mapview::mapview(region)) # not quite what I was after, but a good start

# option 2 - via (new+improved) mapedit
# region_mapedit = mapedit::drawFeatures(map = m)
# sf::write_sf(region_mapedit, "data/scz-mapedit.geojson")
region_mapedit = sf::read_sf("data/scz-mapedit.geojson")
plot(region_mapedit)
region_sp = as(region, "Spatial")
# e = elevatr::get_elev_raster(locations = region_sp, z = 8, src = "aws")
# raster::writeRaster(x = e, filename = "data/scz-elev-z5.tif")
e = raster("data/scz-elev-z5.tif")
x_crs = st_crs(27700)
e_projected = raster::projectRaster(e, crs = x_crs$proj4string)
e_projected

# get points less than 10m above slr
e_mask = e_low = e < 10
plot(e_low)
plot(spData::world, add = T, col = NA)
e_mask[e > 10] = NA
e_low = mask(e, e_mask)
mapview::mapview(e_low)
# out-takes
# e = raster::getData(name = "SRTM", lon = c(-180, 180), lat = c(-60, 60))
# writeRaster(e_low, "scz-elev-low.tif")

system.time({
  r_orig = raster::raster("data/20180816184319_429977190-14-degrees-polish-coast.tif")
  plot(r_orig)
  })
summary(r_orig)
r_agg = aggregate(r_orig, 100)
res(r_agg)
res(e)
plot(r_agg)
plot(e_low, add = T)
e_resampled = raster::resample(e, r_agg)
plot(values(r_agg), values(e_resampled))
cor(values(r_agg), values(e_resampled))^2

# r = extend(r_agg, e)
# r_low = mask(r, e_mask)

# proportion of region that is at risk from slr
e_low_crop = crop(e_low, scz)
plot(e_low_crop)
e_low_sf = spex::polygonize(e_low_crop)
e_low_intersect = st_intersection(e_low_sf, scz)
plot(scz)
plot(e_low_sf, add = T)
plot(e_low_intersect, add = TRUE, col = "blue")
sum(st_area(e_low_intersect)) / st_area(scz)


# publish
scz_buff = st_buffer(scz, dist = 0)
library(tmap)
scz$col = "red"
m = tm_shape(e_low) +
  tm_raster(alpha = 0.5) +
  tm_shape(scz_buff) + # not working...
  tm_fill(col = "red") +
  tm_shape(e_low_intersect) +
  tm_polygons() +
  tm_shape(poland) +
  tm_borders()
m
m2 = mapview::mapview(e_low) + scz
m2
tmap_save(m, "/home/robin/repos/geostats_18/data/scz-map.html")
browseURL("data/scz-map.html")

# elevation data around prague
# prague = osmdata::getbb("prague", format_out = "sf_polygon")
# plot(prague)
# write_sf(prague, "data/prague.geojson")
# prague_elev = get_elev_raster(as(prohonice, "Spatial"), z = 9, src = "aws")
# writeRaster(prague_elev, "data/prague_elev.tif")
prague = read_sf("data/prague.geojson")
prague_elev = raster::raster("data/prague_elev.tif")
plot(prague_elev)
plot(prague, add = TRUE)
# mapview::mapview(prague_elev) +
#   mapview::mapview(prague)

# # detour: finding the resolution:
# p = stplanr::geo_select_aeq(shp = scz_mapedit)
# e_projected = projectRaster(e, crs = p$proj4string)
# summary(e)
# res(e)
# res(e_projected)
# e1 = crop(e, extent(e, 1, 2, 1, 2)) 
# values(e1) = 1:4
# plot(e1)
# e1xy = raster::xyFromCell(e1, cell = 1:ncell(e1))
# e1df = as.data.frame(e1xy)
# e1p = st_as_sf(e1df, coords = c("x", "y"), crs = 4326)
# plot(e1p[1:2, ], cex = 5)
# st_distance(e1p[1, ], e1p[2, ]) # 1.37 km res. (x), 1.49 km res (y) 
# st_distance(e1p[1, ], e1p[3, ]) # 1.37 km res. (x), 1.49 km res (y) 
# 
# # detour: raster -> vector conversion
# e_low_crop = crop(e_low, scz)
# plot(e_low_crop)
# e_low_scz = spPolygons(e_low_crop)
# plot(e_low_scz) # interesting
# class(e_low_scz)
# e_low_sf = st_as_sf(e_low_scz)
# plot(e_low_sf, col = "red") # modern art!

```


## Exercises

1. Use your geocompuational skills to create a reproducible study of the spatial distribution of the impacts of sea level rise, based on assumptions of 2m, 10m and a worst-case 50m scenario of change.
