---
title: "Chapter 12: Transportation"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(spData)
```

# Chapter 12

1) What is the total distance of cycleways that would be constructed if all the routes presented in this [Figure](https://geocompr.robinlovelace.net/transport.html#fig:cycleways) were to be constructed?
    - Bonus: find two ways of arriving at the same answer.

```{r, eval=FALSE}
sum(route_cycleway$distance)
sum(st_length(route_cycleway))
```

2) What proportion of trips represented in the `desire_lines` are accounted for in the `route_cycleway` object?
    - Bonus: what proportion of trips cross the proposed routes?
    - Advanced: write code that would increase this proportion.

```{r, echo=FALSE, eval=FALSE}
sum(route_cycleway$all) / sum(desire_lines$all) # around 2%
d_intersect = desire_lines[route_cycleway, , op = st_crosses]
sum(d_intersect$all) / sum(desire_lines$all) # around 2%
```

3) The analysis presented in this chapter is designed for teaching how geocomputation methods can be applied to transport research. If you were to do this 'for real' for local government or a transport consultancy what top 3 things would you do differently?

Answers:
- Higher level of geographic resolution.
- Use cycle-specific routing services.
- Identify key walking routes.
- Include a higher proportion of trips in the analysis.

4) Clearly the routes identified in this [Figure](https://geocompr.robinlovelace.net/transport.html#fig:cycleways) only provide part of the picture. How would you extend the analysis to incorporate more trips that could potentially be cycled?

5) Imagine that you want to extend the scenario by creating key *areas* (not routes) for investment in place-based cycling policies such as car-free zones, cycle parking points and reduced car parking strategy. How could raster data assist with this work? 
    - Bonus: develop a raster layer that divides the Bristol region into 100 cells (10 by 10) and provide a metric related to transport policy, such as number of people trips that pass through each cell by walking or the average speed limit of roads, from the `bristol_ways` dataset (the approach taken in this [Chapter](https://geocompr.robinlovelace.net/location.html#location)).
