---
title: "Chapter 9: Bridges to GIS software"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(spData)
```

# Chapter 9


1. Create two overlapping polygons (`poly_1` and `poly_2`) with the help of the **sf**-package (see Chapter \@ref(spatial-class)). 

1. Union `poly_1` and `poly_2` using `st_union()` and `qgis:union`.
What is the difference between the two union operations? 
How can we use the **sf** package to obtain the same result as QGIS?

1. Calculate the intersection of `poly_1` and `poly_2` using:

    - **RQGIS**, **RSAGA** and **rgrass7**
    - **sf**

1. Attach `data(dem, package = "RQGIS")` and `data(random_points, package = "RQGIS")`.
Select randomly a point from `random_points` and find all `dem` pixels that can be seen from this point (hint: viewshed).
Visualize your result.
For example, plot a hillshade, and on top of it the digital elevation model, your viewshed output and the point.
Additionally, give `mapview` a try.

1. Use `gdalinfo` via a system call for a raster file stored on disk of your choice.

1. Query all Californian highways from the PostgreSQL/PostGIS database living in the QGIS Cloud introduced in this chapter.
