---
title: "Chapter 9: Bridges to GIS software"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(spData)
```

# Chapter 9

1) Create two overlapping polygons (`poly_1` and `poly_2`) with the help of the **sf**-package (see Chapter 2).

```{r}
# create two polygons
coords_1 =  
  matrix(data = c(0, 0, 1, 0, 1, 1, 0, 1, 0, 0),
         ncol = 2, byrow = TRUE)
coords_2 =
  matrix(data = c(-0.5, -0.5, 0.5, -0.5, 0.5, 0.5, 
                  -0.5, 0.5, -0.5, -0.5),
         ncol = 2, byrow = TRUE)

# create the first polygon
poly_1 = st_polygon(list((coords_1))) 
# convert it into a simple feature collection 
poly_1 = st_sfc(poly_1)
poly_1 = st_sfc(poly_1, crs = 4326)
# finally, convert it into an sf-object
poly_1 = st_sf(geometry = poly_1, data = data.frame(id = 1))

# create a second polygon
poly_2 = st_polygon(list((coords_2))) %>%
  st_sfc(crs = 4326) %>%
  st_sf(data = data.frame(id = 1))
st_geometry(poly_2) = "geometry"
# visualize it
plot(st_geometry(poly_1), xlim = c(-1, 1), ylim = c(-1, 1))
plot(st_geometry(poly_2), add = TRUE)
```

Alternatively:

```{r, eval=FALSE}
centr1 = st_point(c(0, 0))
centr2 = st_point(c(1, 1))
poly_1 = st_buffer(centr1, 1) %>%
  st_sfc(., crs = 4326) %>%
  st_sf(geometry = ., data = data.frame(id = 1))
poly_2 = st_buffer(centr2, 1) %>%
  st_sfc(., crs = 4326) %>%
  st_sf(geometry = ., data = data.frame(id = 1))
# visualize it
plot(st_geometry(poly_1), xlim = c(-1, 2), ylim = c(-1, 2))
plot(st_geometry(poly_2), add = TRUE)
```


2) Union `poly_1` and `poly_2` using `st_union()` and `qgis:union`.
What is the difference between the two union operations? 
How can we use the **sf** package to obtain the same result as QGIS?

```{r, eval=FALSE}
library("RQGIS")
set_env(dev = FALSE)
union_sf = st_union(poly_1, poly_2)
union_qgis = run_qgis("qgis:union", INPUT = poly_1, INPUT2 = poly_2,
                 OUTPUT = file.path(tempdir(), "union.shp"),
                 load_output = TRUE)
par(mfrow = c(1, 2))
plot(st_geometry(union_sf), main = "st_union")
plot(st_geometry(union_qgis), col = 1:3, main = "QGIS union")
# achieve the QGIS result using sf
int = st_intersection(poly_1, poly_2)
# difference_1
diff_1 = st_sym_difference(poly_1, poly_2) %>%
  st_cast("POLYGON")
# the same as:
# diff_2 = st_difference(poly_1, poly_2)
# diff_3 = st_difference(poly_2, poly_1)
# rbind(diff_2, diff_3)
union_sf_2 = rbind(int, diff_1)
plot(st_geometry(union_sf_2), col = 1:3)
```

3) Calculate the intersection of `poly_1` and `poly_2` using:

- **RQGIS**, **RSAGA** and **rgrass7**
- **sf**

```{r, eval=FALSE}
library(sf)
library(RQGIS)

# INTERSECTION USING RQGIS
#***************************
# first of all, we need to find out which function might do this for us
find_algorithms("intersec")
open_help("qgis:intersection")
get_usage("qgis:intersection")
# using R named arguments#
int_qgis = run_qgis("qgis:intersection", INPUT = poly_1, INPUT2 = poly_2,
                    OUTPUT = "int_qgis.shp", load_output = TRUE)
# visualize it
plot(st_geometry(poly_1), xlim = c(-1, 1), ylim = c(-1, 1))
plot(st_geometry(poly_2), add = TRUE)
plot(int_qgis, col = "lightblue", add = TRUE)


# INTERSECTION USING RSAGA
#***************************
# The RSAGA examples only work with SAGA < 2.3. We have informed the
# package maintainer to update SAGA
library(RSAGA)
library(link2GI)
linkSAGA()
rsaga.env()
# save shapefile layers
write_sf(poly_1, file.path(tempdir(), "poly_1.shp"))
write_sf(poly_2, file.path(tempdir(), "poly_2.shp"))
# find out how to union shapefiles with the help of SAGA
rsaga.get.modules(lib = "shapes_polygons")
rsaga.get.usage(lib = "shapes_polygons", module = "Intersect")
# create parameter-argument list for RSAGA
params = list(A = file.path(tempdir(), "poly_1.shp"),
              B = file.path(tempdir(), "poly_2.shp"),
              RESULT = file.path(tempdir(), "int_saga.shp"))
rsaga.geoprocessor(lib = "shapes_polygons", module = "Intersect", 
                   param = params)
int_saga = st_read(file.path(tempdir(), "int_saga.shp"))
# visualize it
plot(st_geometry(poly_1), xlim = c(-1, 1), ylim = c(-1, 1))
plot(st_geometry(poly_2), add = TRUE)
plot(st_geometry(int_saga), col = "lightblue", add = TRUE)

# INTERSECTION USING rgrass7
#***************************
library(link2GI)
library(rgrass7)
link2GI::linkGRASS7(rbind(poly_1, poly_2), ver_select = TRUE)
# let's have a look at the help of v.overlay via rgrass7
execGRASS("g.manual", entry = "v.overlay")
# RQGIS::open_help("grass7:v.overlay")
use_sf()
writeVECT(poly_1, vname = "poly_1")
writeVECT(poly_2, vname = "poly_2")
execGRASS("v.overlay", ainput = "poly_1", binput = "poly_2",
          output = "int_grass", operator = "and", flag = "overwrite")
out_grass = readVECT("int_grass")
plot(st_geometry(poly_1), xlim = c(-1, 1), ylim = c(-1, 1))
plot(st_geometry(poly_2), add = TRUE)
plot(out_grass, add = TRUE, col = "lightblue")

# INTERSECTION USING sf
#***************************

int_sf = st_intersection(poly_1, poly_2)
plot(st_geometry(poly_1), xlim = c(-1, 1), ylim = c(-1, 1))
plot(st_geometry(poly_2), add = TRUE)
plot(int_sf, add = TRUE, col = "lightblue")
```

4) Run `data(dem, package = "RQGIS")` and `data(random_points, package = "RQGIS")`.
Select randomly a point from `random_points` and find all `dem` pixels that can be seen from this point (hint: viewshed).
Visualize your result.
For example, plot a hillshade, and on top of it the digital elevation model, your viewshed output and the point.
Additionally, give `mapview` a try.

```{r, eval=FALSE}
library(RQGIS)
library(raster)
data(dem, package = "RQGIS")
data(random_points, package = "RQGIS")

find_algorithms("viewshed")
alg = "grass7:r.viewshed"
get_usage(alg)
open_help(alg)
# let's find out about the default values
get_args_man(alg)
point = random_points[sample(1:nrow(random_points), 1), ]
coord = paste(sf::st_coordinates(point), collapse = ",")
out = run_qgis(alg, input = dem,
               coordinates = coord,
               output = file.path(tempdir(), "out.tif"),
               load_output = TRUE)

# under Linux this was tested using QGIS 2.18.22 and GRASS GIS 7.4.1. In
# previous versions, however, "grass7:r.viewshed" might not work. In this case
# use rgrass7 directly:
library(rgrass7)
link2GI::linkGRASS7(dem, ver_select = TRUE)
# If both vector and raster are used, until stars raster support is in place,
# use_sp() will be needed. (https://github.com/rsbivand/rgrass7/issues/6)
use_sp()
writeRAST(as(dem, "SpatialGridDataFrame"), "dem")
writeVECT(as(random_points, "Spatial"), vname = "points")
execGRASS("r.viewshed", input = "dem", coordinates = sf::st_coordinates(point),
          output = "view")
out = raster(readRAST("view"))

hs = hillShade(terrain(dem), terrain(dem, "aspect"), 40, 270)
plot(hs, col = gray(0:100 / 100), legend = FALSE)
plot(dem, add = TRUE, alpha = 0.5, legend = FALSE)
plot(point, add = TRUE, col = "red", pch = 16)
plot(out, add = TRUE, col = "lightgray", legend = FALSE)
plot(point, add = TRUE, col = "red", pch = 16)

# or using mapview
library(mapview)
mapview(out, col = "white", map.type = "Esri.WorldImagery") +
  mapview(point)
```

5) Compute catchment area and catchment slope of `data("dem", package = "RQGIS")` using **RSAGA**.

```{r, eval=FALSE}
library("RSAGA")
# check if rsaga.env can find the corresponding SAGA installation
rsaga.env()
rsaga.get.libraries()
rsaga.get.modules(libs = "ta_hydrology")
rsaga.get.usage(lib = "ta_hydrology", module = "SAGA Wetness Index")
# RSAGA only accepts files living on disk, so we have to save dem 
raster::writeRaster(dem, filename = file.path(tempdir(), "dem.sdat"),
                    format = "SAGA")
# construct a parameter-argument list
params = list(DEM = file.path(tempdir(), "dem.sgrd"),
              AREA = file.path(tempdir(), "carea.sdat"),
              SLOPE = file.path(tempdir(), "cslope.sdata"),
              AREA_TYPE = 0,
              SLOPE_TYPE = 1)
rsaga.geoprocessor(lib = "ta_hydrology", module = "SAGA Wetness Index",
                   param = params)
# load the output back into R again
out = stack(file.path(tempdir(), "carea.sdat"), 
            file.path(tempdir(), "cslope.sdat"))
plot(out)
```

6) Use `gdalinfo` via a system call for a raster file stored on disk of your choice.

```{r, eval=FALSE}
link2GI::linkGDAL()
data("dem", package = "RQGIS")
file = file.path(tempdir(), "dem.tif")
writeRaster(dem, filename = file, prj = TRUE, format = "GTiff")
cmd = paste("gdalinfo", file)
system(cmd)
```

7) Query all Californian highways from the PostgreSQL/PostGIS database living in the QGIS Cloud introduced in this chapter.

```{r, eval=FALSE}
library(RPostgreSQL)
library(sf)
conn = dbConnect(drv = PostgreSQL(), dbname = "rtafdf_zljbqm",
                 host = "db.qgiscloud.com",
                 port = "5432", user = "rtafdf_zljbqm", 
                 password = "d3290ead")
query = paste(
  "SELECT *",
  "FROM highways",
  "WHERE state = 'CA';")
ca_highways = st_read(conn, query = query, geom = "wkb_geometry")
plot(st_geometry(ca_highways))
```
