---
title: "Chapter 3: Attribute operations"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(raster)
library(tidyverse)
```

For these exercises we will use the `us_states` and `us_states_df` datasets from the **spData** package:

```{r}
library(spData)
data(us_states)
data(us_states_df)
```

`us_states` is a spatial object (of class `sf`), containing geometry and a few attributes (including name, region, area, and population) of states within the contiguous United States.
`us_states_df` is a data frame (of class `data.frame`) containing the name and additional variables (including median income and poverty level, for years 2010 and 2015) of US states, including Alaska, Hawaii and Puerto Rico.
The data comes from the US Census Bureau, and is documented in `?us_states` and `?us_states_df`.

# Chapter 3

1) Create a new object called `us_states_name` that contains only the `NAME` column from the `us_states` object. 
What is the class of the new object and what makes it geographic? 

```{r}
us_states_name = us_states %>% dplyr::select(NAME)
class(us_states_name)
```

  - It is of class `sf` and `data.frame`: it has 2 classes.
  - It is the `sf` class that makes in geographic.
  - More specifically it is the attributes of the object (`sf_column`) and the geometry column (such as `bbox`, `crs`) that make it geographic.

```{r}
attributes(us_states_name)
attributes(us_states_name$geometry)
```

2) Select columns from the `us_states` object which contain population data.
Obtain the same result using a different command (bonus: try to find three ways of obtaining the same result).
Hint: try to use helper functions, such as `contains` or `starts_with` from **dplyr** (see `?contains`).

```{r}
us_states %>% dplyr::select(total_pop_10, total_pop_15)

# or
us_states %>% dplyr::select(starts_with("total_pop"))

# or
us_states %>% dplyr::select(contains("total_pop"))
```

3) Find all states with the following characteristics (bonus find *and* plot them):
    - Belong to the Midwest region.
```{r}
us_states %>% 
  filter(REGION == "Midwest")
```
    - Belong to the West region, have an area below 250,000 km^2^ *and* in 2015 a population greater than 5,000,000 residents (hint: you may need to use the function `units::set_units()` or `as.numeric()`).
```{r}
us_states %>% filter(REGION == "West", AREA < units::set_units(250000, km^2),total_pop_15 > 5000000)
# or
us_states %>% filter(REGION == "West", as.numeric(AREA) < 250000,total_pop_15 > 5000000)
```
    - Belong to the South region, had an area larger than 150,000 km^2^ or a total population in 2015 larger than 7,000,000 residents.
```{r}
us_states %>% filter(REGION == "South", AREA > units::set_units(150000, km^2), total_pop_15 > 7000000)
# or
us_states %>% filter(REGION == "South", as.numeric(AREA) > 150000, total_pop_15 > 7000000)
```

4) What was the total population in 2015 in the `us_states` dataset?
What was the minimum and maximum total population in 2015?
```{r}
us_states %>% summarize(total_pop = sum(total_pop_15),
                        min_pop = min(total_pop_15),
                        max_pop = max(total_pop_15))
```

5) How many states are there in each region?
```{r}
us_states %>%
  group_by(REGION) %>%
  summarize(nr_of_states = n())
```

6) What was the minimum and maximum total population in 2015 in each region?
What was the total population in 2015 in each region?
```{r}
us_states %>%
  group_by(REGION) %>%
  summarize(min_pop = min(total_pop_15),
            max_pop = max(total_pop_15),
            tot_pop = sum(total_pop_15))
```

7) Add variables from `us_states_df` to `us_states`, and create a new object called `us_states_stats`.
What function did you use and why?
Which variable is the key in both datasets?
What is the class of the new object?
```{r}
us_states_stats = us_states %>%
  left_join(us_states_df, by = c("NAME" = "state"))
class(us_states_stats)
```

8) `us_states_df` has two more variables than `us_states`.
How can you find them? (hint: try to use the `dplyr::anti_join()` function)
```{r}
us_states_df %>%
  anti_join(us_states, by = c("state" = "NAME"))
```

9) What was the population density in 2015 in each state?
What was the population density in 2010 in each state?
```{r}
us_states2 = us_states %>%
  mutate(pop_dens_15 = total_pop_15/AREA,
         pop_dens_10 = total_pop_10/AREA)
```

10) How much has population density changed between 2010 and 2015 in each state?
Calculate the change in percentages and map them.
```{r}
us_popdens_change = us_states2 %>%
  mutate(pop_dens_diff_10_15 = pop_dens_15 - pop_dens_10,
         pop_dens_diff_10_15p = (pop_dens_diff_10_15/pop_dens_15) * 100)
plot(us_popdens_change["pop_dens_diff_10_15p"])
```

11) Change the columns names in `us_states` to lowercase. (Hint: helper functions - `tolower()` and `colnames()` may help).
```{r}
us_states %>%
  set_names(tolower(colnames(.)))
```

12) Using `us_states` and `us_states_df` create a new object called `us_states_sel`.
The new object should have only two variables - `median_income_15` and `geometry`.
Change the name of the `median_income_15` column to `Income`.
```{r}
us_states_sel = us_states %>%
  left_join(us_states_df, by = c("NAME" = "state")) %>%
  dplyr::select(Income = median_income_15)
```

13) Calculate the change in median income between 2010 and 2015 for each state.
Bonus: what was the minimum, average and maximum median income in 2015 for each region?
What is the region with the largest increase of the median income?
```{r}
us_income_change = us_states %>%
  left_join(us_states_df, by = c("NAME" = "state")) %>%
  mutate(income_change = median_income_15 - median_income_10) 

us_income_change_reg = us_income_change %>%
  group_by(REGION) %>%
  summarize(min_income_change = min(income_change),
            mean_income_change = mean(income_change),
            max_income_change = max(income_change))

us_income_change_reg %>%
  filter(mean_income_change == max(mean_income_change)) %>%
  pull(REGION) %>%
  as.character()
```

14) Create a raster from scratch with nine rows and columns and a resolution of 0.5 decimal degrees (WGS84).
Fill it with random numbers.
Extract the values of the four corner cells. 
```{r}
r = raster(nrow = 9, ncol = 9, res = 0.5, xmn = 0, xmx = 4.5,
           ymn = 0, ymx = 4.5, vals = rnorm(81))
# using cell IDs
r[c(1, 9, 81 - 9, 81)]
# using indexing
r[c(1, nrow(r)), c(1, ncol(r))]
# corresponds to [1, 1], [1, 9], [9, 1], [9, 9]
```


15) What is the most common class of our example raster `grain` (hint: `modal()`)?
```{r}
grain_size = c("clay", "silt", "sand")
grain = raster(nrow = 6, ncol = 6, res = 0.5, 
               xmn = -1.5, xmx = 1.5, ymn = -1.5, ymx = 1.5,
               vals = factor(sample(grain_size, 36, replace = TRUE), 
                             levels = grain_size))
cellStats(grain, modal) %>%
  factorValues(grain, .)
factorValues(grain, modal(values(grain)))
```

16) Plot the histogram and the boxplot of the `data(dem, package = "RQGIS")` raster. 
```{r}
data(dem, package = "RQGIS")
par(mfrow = c(1, 2))
hist(dem)
boxplot(dem)
```

17) Now attach also `data(ndvi, package = "RQGIS")`. 
Create a raster stack using `dem` and `ndvi`, and make a `pairs()` plot.
```{r}
data(ndvi, package = "RQGIS")
s = stack(dem, ndvi)
pairs(s)
```
