---
title: "Chapter 1"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(spData)
```

# Chapter 1

1. Think about the terms 'GIS', 'GDS' and 'Geocomputation' described above. Which is your favorite and and why?

1. Provide 3 reasons for using a scriptable language such R for geocomputation instead of established GIS programs such as QGIS.

- Reproducibility: a sequence of operations is much easier to save and share when using a scripting language.
- Scalability: scripting languages make it easy to break-up the work and parallelise, making it more scalable.
- Flexibility: the building blocks of the language allows it to interface efficiently with external software, enabling R to perform a wide range of tasks (e.g. animations, online interactive maps).
- Efficiency: the GUI of GIS programs can be slow to use and install.

1. Name two advantages and two disadvantages of using the older **sp** package compared with the new **sf** package.

- `+` Number of packages supported.
- `+` Stability
- `-` Slow performance
- `-` Non standard data format
