---
title: "Chapter 10"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
# source("https://github.com/Robinlovelace/geocompr/raw/master/code/chapters/10-algorithms.R")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(spData)
```

# Chapter 10


1. Read the script `10-centroid-alg.R` in the `code` folder of the book's GitHub repo.
    - Which of the best practices covered in section \@ref(scripts) does it follow?
    - Create a version of the script on your computer in an IDE such as RStudio (preferably by typing-out the script line-by-line, in your own coding style and with your own comments, rather than copy-pasting --- this will help you learn how to type scripts) and, using the example of a square polygon (e.g. created with `poly_mat = cbind(x = c(0, 0, 9, 9, 0), y = c(0, 9, 9, 0, 0))`) execute the script line-by-line.
    - What changes could be made to the script to make it more reproducible?
    <!-- - Answer: The script could state that it needs a an object called `poly_mat` to be present and, if none is present, create an example dataset at the outset for testing. -->
<!-- 1. Check-out the script `10-earthquakes.R` in the `code` folder of the book's GitHub [repo](https://github.com/Robinlovelace/geocompr/blob/master/code/10-earthquakes.R). -->
<!--     - Try to reproduce the results: how many significant earthquakes were there last month? -->
<!--     - Modify the script so that it provides a map with all earthquakes that happened in the past hour. -->
<!-- change line 10 to: -->
<!-- u = "https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/significant_hour.geojson" -->
    - How could the documentation be improved?
  <!-- It could document the source of the data better - e.g. with `data from https://earthquake.usgs.gov/earthquakes/feed/v1.0/geojson.php` -->
1. In section \@ref(geographic-algorithms) we calculated the area and geographic centroid of the polygon represented by `poly_mat`.
    - Reproduce the results on your own computer with reference to the script `10-centroid-alg.R`, an implementation of this algorithm (bonus: type out the commands - try to avoid copy-pasting).
    <!-- Todo: add link to that script file (RL) -->
    - Are the results correct? Verify them by converting `poly_mat` into an `sfc` object (named `poly_sfc`) with `st_polygon()` (hint: this function takes objects of class `list()`) and then using `st_area()` and `st_centroid()`.
<!-- We can verify the answer by converting `poly_mat` into a simple feature collection as follows, which shows the calculations match: -->
```{r, eval=FALSE}
# FIX NEEDED
poly_sfc = sf::st_polygon(list(poly_mat))
sf::st_area(poly_sfc)
sf::st_centroid(poly_sfc)
```
1. It was stated that the algorithm we created only works for *convex hulls*. Define convex hulls (see Chapter \@ref(geometric-operations)) and test the algorithm on a polygon that is *not* a convex hull.
     - Bonus 1: Think about why the method only works for convex hulls and note changes that would need to be made to the algorithm for other types of shape to be calculated.
<!-- The algorithm would need to be able to have negative as well as positive area values. -->
     - Bonus 2: Building on the contents of `10-centroid-alg.R`, write an algorithm only using base R functions that can find the total length of linestrings represented in matrix form.
<!-- Todo: add example of matrix representing a linestring, demonstrate code to verify the answer, suggest alternative functions to decompose as a bonus. -->
1. In section \@ref(functions) we created a different versions of the `poly_centroid()` function that generated outputs of class `sfg` (`poly_centroid_sfg()`) and type-stable `matrix` outputs (`poly_centroid_type_stable()`). Further extend the function by creating a version (e.g. called `poly_centroid_sf()`) that is type stable (only accepts inputs of class `sf`) *and* returns `sf` objects (hint: you may need to convert the object `x` into a matrix with the command `sf::st_coordinates(x)`.
    - Verify it works by running `poly_centroid_sf(sf::st_sf(sf::st_sfc(poly_sfc)))`
    - What error message do you get when you try to run `poly_centroid_sf(poly_mat)`?
    
```{r, eval=FALSE}
# FIX NEEDED
poly_centroid_sf = function(x) {
  stopifnot(is(x, "sf"))
  xcoords = sf::st_coordinates(x)
  centroid_coords = poly_centroid(xcoords)
  centroid_sf = sf::st_sf(geometry = sf::st_sfc(sf::st_point(centroid_coords)))
  centroid_sf
}
poly_centroid_sf(sf::st_sf(sf::st_sfc(poly_sfc)))
poly_centroid_sf(poly_sfc)
poly_centroid_sf(poly_mat)
```
