---
title: "Chapter 14"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(spData)
```

# Chapter 14


1. Run a NMDS using the percentage data of the community matrix. 
Report the stress value and compare it to the stress value as retrieved from the NMDS using presence-absence data.
What might explain the observed difference?
1. Compute catchment area and catchment slope using **RSAGA** (see section \@ref(rsaga)).
1. Use profile and tangential curvature as additional predictors for the spatial prediction of the floristic gradient (hint: `grass7:r.slope.aspect`).
1. Retrieve the bias-reduced RMSE using spatial cross-validation including the estimation of optimal hyperparameter combinations (random search with 50 iterations) in an inner tuning loop (see section \@ref(svm)).
Parallelize the tuning level (see section \@ref(svm)).
Report the mean RMSE and use a boxplot to visualize all retrieved RMSEs.
