---
title: "Chapter 4"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(spData)
```

# Chapter 4

4. Use `data(dem, package = "RQGIS")`, and reclassify the elevation in three classes: low, middle and high.
Secondly, compute the NDVI (`data(ndvi, package = "RQGIS")`) and the mean elevation for each altitudinal class.
```{r, eval=FALSE}
library(classInt)
data(dem, package = "RQGIS")
summary(dem)
# find quantile breaks
brk = classIntervals(values(dem), n = 3)$brk
# also try
# breask = classIntervals(values(dem), n = 3, style = "fisher")
# construct reclassification matrix
rcl = matrix(c(brk[1] - 1, brk[2], 1, brk[2], brk[3], 2, brk[3], brk[4], 3), 
             ncol = 3, byrow = TRUE)
# reclassify altitudinal raster
recl = reclassify(dem, rcl = rcl)
# compute the mean dem and ndvi values for each class
zonal(stack(dem, ndvi), recl, fun = "mean")
```

5. Apply a line detection filter to `data(dem, package = "RQGIS")`.
```{r, eval=FALSE}
# from the focal help page:
# Laplacian filter: filter=matrix(c(0,1,0,1,-4,1,0,1,0), nrow=3)
# Sobel filter: filter=matrix(c(1,2,1,0,0,0,-1,-2,-1) / 4, nrow=3)
# compute the Sobel filter
# check if there are NAs
is.na(dem)  # just 0s, so no NAs
sobel = focal(dem, w = matrix(c(1, 2, 1, 0, 0, 0, -1, -2, -1) / 4, nrow = 3))
# CHECK IF THIS IS CORRECT
```

6. Calculate the NDVI of a Landsat image. 
Use the Landsat image provided by the **spDataLarge** package (`system.file("raster/landsat.tif", package="spDataLarge")`).

```{r, eval=FALSE}
file = system.file("raster/landsat.tif", package="spDataLarge")
r = stack(file)
# compute NDVI manually
ndvi = (r[["landsat.4"]] - r[["landsat.3"]]) / (r[["landsat.4"]] + r[["landsat.3"]])
# compute NDVI with the help of RStoolbox
library(RStoolbox)
ndvi_rstoolbox = spectralIndices(r, red = 3, nir = 4, indices = "NDVI")
all.equal(ndvi, ndvi_rstoolbox)
```

7. This [post](https://stackoverflow.com/questions/35555709/global-raster-of-geographic-distances) shows how to use `raster::distance()`.
Extract Spain, calculate a distance raster and weight it with elevation.
Finally, compute the difference between the raster using the euclidean distance and the raster weighted by elevation.
(Hint: Have a look at `getData()` to retrieve a digital elevation model for Spain.)

```{r, eval=FALSE}
library(raster)
# find out the ISO_3 code of Spain
dplyr::filter(ccodes(), NAME %in% "Spain")
# retrieve a dem of Spain
dem = getData("alt", country = "ESP", mask = FALSE)
# change the resolution to decrease computing time
agg = aggregate(dem, fact = 5)
poly = getData("GADM", country = "ESP", level = 1)
plot(dem)
plot(poly, add = TRUE)
# visualize NAs
plot(is.na(agg))
# construct a distance input raster
# we have to set the land cells to NA and the sea cells to an arbitrary value since 
# raster::distance computes the distance to the nearest non-NA cell
dist = is.na(agg)
cellStats(dist, summary)
# convert land cells into NAs and sea cells into 1s
dist[dist == FALSE] = NA
dist[dist == TRUE] = 1
plot(dist)
# compute distance to nearest non-NA cell
dist = raster::distance(dist)
# just keep Spain
dist = mask(dist, poly)
# convert distance into km
dist = dist / 1000
# now let's weight each 100 altitudinal meters by an additionaly distance of 10 km
agg = mask(agg, poly)
agg[agg < 0] = 0
weight = dist + agg / 100 * 10
plot(weight - dist)
```
