---
title: "Chapter 5: Geometric operations"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(raster)
library(tidyverse)
library(spData)
library(spDataLarge)
library(RQGIS)
```

Some of the exercises use a vector (`random_points`) and raster dataset (`ndvi`) from the **RQGIS** package.
They also use a polygonal 'convex hull' derived from the vector dataset (`ch`) to represent the area of interest:
```{r, message=FALSE}
data(random_points)
data(ndvi)
ch = st_combine(random_points) %>% 
  st_convex_hull()
```

# Chapter 5

1. Generate and plot simplified versions of the `nz` dataset.
Experiment with different values of `keep` (ranging from 0.5 to 0.00005) for `ms_simplify()` and `dTolerance` (from 100 to 100,000) `st_simplify()` .
    - At what value does the form of the result start to break-down for each method, making New Zealand unrecognizable?
    - Advanced: What is different about the geometry type of the results from `st_simplify()` compared with the geometry type of `ms_simplify()`? 
What problems does this create and how can this be resolved?
```{r}
plot(rmapshaper::ms_simplify(st_geometry(nz), keep = 0.5))
plot(rmapshaper::ms_simplify(st_geometry(nz), keep = 0.05))
# Starts to breakdown here at 0.5% of the points:
plot(rmapshaper::ms_simplify(st_geometry(nz), keep = 0.005))
# At this point no further simplification changes the result
plot(rmapshaper::ms_simplify(st_geometry(nz), keep = 0.0005))
plot(rmapshaper::ms_simplify(st_geometry(nz), keep = 0.00005))
plot(st_simplify(st_geometry(nz), dTolerance = 100))
plot(st_simplify(st_geometry(nz), dTolerance = 1000))
# Starts to breakdown at 10 km:
plot(st_simplify(st_geometry(nz), dTolerance = 10000))
plot(st_simplify(st_geometry(nz), dTolerance = 100000))
plot(st_simplify(st_geometry(nz), dTolerance = 100000, preserveTopology = TRUE))

# Problem: st_simplify returns POLYGON and MULTIPOLYGON results, affecting plotting
# Cast into a single geometry type to resolve this
nz_simple_poly = st_simplify(st_geometry(nz), dTolerance = 10000) %>% 
  st_sfc() %>% 
  st_cast("POLYGON")
nz_simple_multipoly = st_simplify(st_geometry(nz), dTolerance = 10000) %>% 
  st_sfc() %>% 
  st_cast("MULTIPOLYGON")
plot(nz_simple_poly)
length(nz_simple_poly)
nrow(nz)
```
2. In the first exercise in this [Chapter](https://geocompr.robinlovelace.net/spatial-data-operations.html) it was established that Canterbury region had 70 of the 101 highest points in New Zealand.
Using `st_buffer()`, how many points in `nz_height` are within 100 km of Canterbury?
```{r}
canterbury = nz[nz$Name == "Canterbury", ]
cant_buff = st_buffer(canterbury, 100)
nz_height_near_cant = nz_height[cant_buff, ]
nrow(nz_height_near_cant) # 75 - 5 more
```
3. Find the geographic centroid of New Zealand.
How far is it from the geographic centroid of Canterbury?
```{r}
cant_cent = st_centroid(canterbury)
nz_centre = st_centroid(st_union(nz))
st_distance(cant_cent, nz_centre) # 234 km
```
4. Most world maps have a north-up orientation.
A world map with a south-up orientation could be created by a reflection (one of the affine transformations not mentioned in this [section](https://geocompr.robinlovelace.net/geometric-operations.html#affine-transformations)) of the `world` object's geometry.
Write code to do so.
Hint: you need to use a two-element vector for this transformation.
    - Bonus: create a upside down map of your country.
```{r}
world_sfc = st_geometry(world)
world_sfc_mirror = world_sfc * c(1, -1)
plot(world_sfc)
plot(world_sfc_mirror)

us_states_sfc = st_geometry(us_states)
us_states_sfc_mirror = us_states_sfc * c(1, -1)
plot(us_states_sfc)
plot(us_states_sfc_mirror)
## nicer plot
# library(ggrepel)
# us_states_sfc_mirror_labels = st_centroid(us_states_sfc_mirror) %>% 
#   st_coordinates() %>%
#   as_data_frame() %>% 
#   mutate(name = us_states$NAME)
# us_states_sfc_mirror_sf = st_set_geometry(us_states, us_states_sfc_mirror)
# ggplot(data = us_states_sfc_mirror_sf) +
#   geom_sf(color = "white") +
#   geom_text_repel(data = us_states_sfc_mirror_labels, mapping = aes(X, Y, label = name), size = 3, min.segment.length = 0) +
#   theme_void() 
```
5. Subset the point in `p` that is contained within `x` *and* `y` (see this [section](https://geocompr.robinlovelace.net/geometric-operations.html#clipping) and this [Figure](https://geocompr.robinlovelace.net/geometric-operations.html#fig:venn-clip)).
    - Using base subsetting operators.
    - Using an intermediary object created with `st_intersection()`.
```{r}
# data preparation
b = st_sfc(st_point(c(0, 1)), st_point(c(1, 1))) # create 2 points
b = st_buffer(b, dist = 1) # convert points to circles
x = b[1]
y = b[2]
bb = st_bbox(st_union(x, y))
pmat = matrix(c(bb[c(1, 2, 3, 2, 3, 4, 1, 4, 1, 2)]), ncol = 2, byrow = TRUE)
box = st_polygon(list(pmat))
set.seed(2017)
p = st_sample(x = box, size = 10)
```
```{r}
p_in_y = p[y]
p_in_xy = p_in_y[x]
x_and_y = st_intersection(x, y)
p[x_and_y]
```
6. Calculate the length of the boundary lines of US states in meters.
Which state has the longest border and which has the shortest?
Hint: The `st_length` function computes the length of a `LINESTRING` or `MULTILINESTRING` geometry.
```{r}
us_states2163 = st_transform(us_states, 2163)
us_states_bor = st_cast(us_states2163, "MULTILINESTRING")
us_states_bor$borders = st_length(us_states_bor)
arrange(us_states_bor, borders) %>% slice(1)
arrange(us_states_bor, -borders) %>% slice(1)
```
7. Crop the `ndvi` raster using (1) the `random_points` dataset and (2) the `ch` dataset.
Are there any difference in the output maps?
Next, mask `ndvi` using these two datasets.
Can you see any difference now?
How can you explain that?
```{r}
plot(ndvi)
plot(st_geometry(random_points), add = TRUE)
plot(ch, add = TRUE)

ndvi_crop1 = crop(ndvi, as(random_points, "Spatial")) 
ndvi_crop2 = crop(ndvi, as(ch, "Spatial")) 
plot(ndvi_crop1)
plot(ndvi_crop2)

ndvi_mask1 = mask(ndvi, as(random_points, "Spatial")) 
ndvi_mask2 = mask(ndvi, as(ch, "Spatial")) 
plot(ndvi_mask1)
plot(ndvi_mask2)
```
8. Firstly, extract values from `ndvi` at the points represented in `random_points`.
Next, extract average values of `ndvi` using a 90 buffer around each point from `random_points` and compare these two sets of values. 
When would extracting values by buffers be more suitable than by points alone?
```{r}
random_points_buf = st_buffer(random_points, dist = 90)
plot(ndvi)
plot(st_geometry(random_points_buf), add = TRUE)
plot(ch, add = TRUE)
random_points$ndvi = raster::extract(ndvi, as(random_points, "Spatial"), buffer = 90, fun = mean)
random_points$ndvi2 = raster::extract(ndvi, as(random_points, "Spatial"))
plot(random_points$ndvi, random_points$ndvi2)
```
9. Subset points higher than 3100 meters in New Zealand (the `nz_height` object) and create a template raster with a resolution of 3km. 
Using these objects:
    - Count numbers of the highest points in each grid cell.
    - Find the maximum elevation in each grid cell.
```{r}
nz_height3100 = dplyr::filter(nz_height, elevation > 3100)
new_graticule = st_graticule(nz_height3100, datum = 2193)
plot(st_geometry(nz_height3100), graticule = new_graticule, axes = TRUE)
nz_template = raster(extent(nz_height3100), resolution = 3000,
                         crs = st_crs(nz_height3100)$proj4string)
nz_raster = rasterize(nz_height3100, nz_template, 
                       field = "elevation", fun = "count")
plot(nz_raster)
nz_raster2 = rasterize(nz_height3100, nz_template, 
                       field = "elevation", fun = max)
plot(nz_raster2)
```

10. Aggregate the raster counting high points in New Zealand (created in the previous exercise), reduce its geographic resolution by half (so cells are 6 by 6 km) and plot the result.
    - Resample the lower resolution raster back to a resolution of 3 km. How have the results changed?
    - Name two advantages and disadvantages of reducing raster resolution.
```{r, eval=FALSE}
nz_raster_low = raster::aggregate(nz_raster, fact = 2, fun = sum)
res(nz_raster_low)
nz_resample = resample(nz_raster_low, nz_raster)
plot(nz_raster_low)
plot(nz_resample) # the results are spread over a greater area and there are border issues
plot(nz_raster)
```

advantage: lower memory use
advantage: faster processing
advantage: good for viz in some cases
disadvantage: removes geographic detail
disadvantage: another processing step

11. Polygonize the `grain` dataset and filter all squares representing clay.
    - Name two advantages and disadvantages of vector data over raster data.
    - At which points would it be useful to convert rasters to vectors in your work?

```{r}
grain_poly = rasterToPolygons(grain) %>% 
  st_as_sf()
levels(grain)
clay = dplyr::filter(grain_poly, layer == 1)
plot(clay)
```

advantages:
- can be used to subset other vector objects
- can do affine transformations and use sf/dplyr verbs
disadvantages:
- better consistency, fast processing on some operations, functions developed for some domains
