---
title: "Chapter 13: Geomarketing"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(raster)
library(tidyverse)
library(spData)
```

```{r}
data("census_de", package = "spDataLarge")
input = dplyr::select(census_de, x = x_mp_1km, y = y_mp_1km, pop = Einwohner,
                      women = Frauen_A, mean_age = Alter_D,
                      hh_size = HHGroesse_D)
input_tidy = mutate_all(input, funs(ifelse(. %in% c(-1, -9), NA, .)))
```


# Chapter 13

1) We have used `raster::rasterFromXYZ()` to convert a `input_tidy` into a raster brick. 
Try to achieve the same with the help of the `sp::gridded()` function.

```{r}
input = st_as_sf(input, coords = c("x", "y"))
# use the correct projection (see data description)
input = st_set_crs(input, 3035)
# convert into an sp-object
input = as(input, "Spatial")
gridded(input) = TRUE
# convert into a raster stack
input = stack(input)
```

2) Download the csv file containing inhabitant information for a 100 m cell resolution (https://www.zensus2011.de/SharedDocs/Downloads/DE/Pressemitteilung/DemografischeGrunddaten/csv_Bevoelkerung_100m_Gitter.zip?__blob=publicationFile&v=3).
Please note that the unzipped file has a size of 1.23 GB.
To read it into R you can use `readr::read_csv`.
This takes 30 seconds on my machine (16 GB RAM)
`data.table::fread()` might be even faster, and returns an object of class `data.table()`.
Use `as.tibble()` to convert it into a tibble.
Build an inhabitant raster, aggregate it to a cell resolution of 1 km, and compare the difference with the inhabitant raster (`inh`) we have created using class mean values.

```{r, eval=FALSE}
build_census_raster = function(url) {
  download.file(url = url, destfile = file.path(tempdir(), "census.zip"),
                method = "auto", mode = "wb")
  # list the file names
  nms = unzip(file.path(tempdir(), "census.zip"), list = TRUE)
  # unzip only the csv file
  base_name = grep(".csv$", nms$Name, value = TRUE)
  unzip(file.path(tempdir(), "census.zip"), files = base_name, exdir = tempdir())
  # read in the csv file
  input = data.table::fread(file.path(tempdir(), base_name)) %>%
    as.tibble
  input = dplyr::select(input, x = starts_with("x_mp_1"), y = starts_with("y_mp_1"),
                        inh = Einwohner)
  # set -1 and -9 to NA
  input = mutate_all(input, funs(ifelse(. %in% c(-1, -9), NA, .)))
  # convert table into a raster (x and y are cell midpoints)
  coordinates(input) =~ x + y
  # use the correct projection
  proj4string(input) = CRS("+init=epsg:3035")
  gridded(input) = TRUE
  # convert into a raster stack
  raster(input)
}

# download 1km resolution
url = paste0("https://www.zensus2011.de/SharedDocs/Downloads/DE/", 
             "Pressemitteilung/DemografischeGrunddaten/csv_Zensusatlas_", 
             "klassierte_Werte_1km_Gitter.zip?__blob=publicationFile&v=8")
inp_coarse = build_census_raster(url)
# reclassify
rcl = matrix(c(1, 1, 125, 2, 2, 375, 3, 3, 1250, 4, 4, 3000, 5, 5, 6000,
               6, 6, 8000), ncol = 3, byrow = TRUE)
inh_coarse = reclassify(inp_coarse, rcl = rcl, right = NA)

# Download and build 1km inhabitant raster
url = paste0("https://www.zensus2011.de/SharedDocs/Downloads/DE/Pressemitteilung/",
             "DemografischeGrunddaten/csv_Bevoelkerung_100m_Gitter.zip", 
             "?__blob=publicationFile&v=3")
inh_fine = build_census_raster(url)
inh_fine = aggregate(inh_fine, fact = 1000 / res(inp_fine)[1], fun = sum)
inh_fine - inh_coarse  # origin has to be the same
origin(inh_fine) = origin(inh_coarse)
summary(inh_fine - inh_coarse)
plot(inh_fine - inh_coarse)
plot(abs(inh_fine - inh_coarse) > 1000)
cellStats((abs(inh_fine - inh_coarse) > 1000), stat = "sum")
cellStats((abs(inh_fine - inh_coarse) > 5000), stat = "sum")
```

3) Suppose our bike shop predominantly sold electric bikes to older people. 
Change the age raster accordingly, repeat the remaining analyses and compare the changes with our original result.
