---
title: "Chapter 7: Geographic data I/O"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(raster)
library(tidyverse)
library(spData)
```

# Chapter 7

1. List and describe three types of vector, raster, and geodatabase formats.

2. Name at least two differences between `read_sf()` and the more well-known function `st_read()`.

3. Read the `cycle_hire_xy.csv` file from the **spData** package (Hint: it is located in the `misc\` folder).
What is a geometry type of the loaded object? 

4. Download the borders of Germany using **rnaturalearth**, and create a new object called `germany_borders`.
Write this new object to a file of the GeoPackage format.

5. Download the global monthly minimum temperature with a spatial resolution of five minutes using the **raster** package.
Extract the June values, and save them to a file named `tmin_june.tif` file (hint: use `raster::subset()`).

6. Create a static map of Germany's borders, and save it to a PNG file.

7. Create an interactive map using data from the `cycle_hire_xy.csv` file. 
Export this map to a file called `cycle_hire.html`.
