---
title: "Chapter 8"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(spData)
library(tmap)
```

# Chapter 8


For these exercises we will create a new object, `africa`, using the `world` and `worldbank_df` datasets from the **spData** package (see chapter \@ref(attr) to learn more about attribute operations) as follows:

```{r}
africa = world %>% 
  filter(continent == "Africa", !is.na(iso_a2)) %>% 
  left_join(worldbank_df, by = "iso_a2") %>% 
  dplyr::select(name, subregion, gdpPercap, HDI, pop_growth) %>% 
  st_transform("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25")
```

We will also use `zion` and `nlcd` datasets from **spDataLarge**:

```{r, results='hide'}
zion = st_read((system.file("vector/zion.gpkg", package = "spDataLarge")))
```

1. Create a map showing the geographic distribution of the Human Development Index (`HDI`) across Africa with 
-base graphics (hint: use `plot()`) and
-**tmap** (hint: use `tm_shape(africa) + ...`).
    - Name two advantages of each approach
    - What three other mapping packages could be used to show the same data?
    - Name an advantage of each
    - Bonus: create three more maps of Africa using the three packages mentioned previously
1. Extend the map of Africa created with **tmap** for the previous exercise so the legend has three bins: "High" (`HDI` above 0.7), "Medium" (`HDI` between 0.55 and 0.7) and "Low" (`HDI` below 0.55).
    - Bonus: improve the map aesthetics, for example by changing the legend title, class labels and color palette.
```{r}
# >0.8 - Very high human development
# >0.7 - High human development
# >0.55 - Medium human development
# <0.55 - Low human development
sm1 = tm_shape(africa) +
  tm_polygons(
  col = "HDI",
  title = "Human Development Index",
  breaks = c(0, 0.55, 0.7, 0.8),
  labels = c("Low", "Medium", "High"),
  palette = "YlGn"
  )
```
1. Represent `africa`'s subregions on the map. 
Change the default color palette and legend title.
Next, combine this map and the map created in the previous exercise into a single plot.
```{r}
sm2 = tm_shape(africa) +
  tm_polygons(col = "subregion",
              title = "Subregion", 
              palette = "Set2")
tmap_arrange(sm2, sm1)
```
1. Create a land cover map of the Zion National Park.
    - Change the default colors to match your perception of the land cover categories
    - Add a scale bar and north arrow and change the position of both to improve the map's aesthetic appeal
    - Bonus: Add an inset map of the Zion National Park's location in the context of the Utah state. (Hint: an object representing Utah can be subsetted from the `us_states` dataset.) 
```{r, eval=FALSE}
# FIX NEEDED
lc_colors = c("#476ba0", "#aa0000", "#b2ada3", "#68aa63", "#a58c30",
              "#c9c977", "#dbd83d", "#bad8ea")
sm3 = tm_shape(nlcd) + tm_raster(palette = lc_colors, title = "Land cover") +
  tm_shape(zion) + tm_borders(lwd = 3) +
  tm_scale_bar(size = 1, position = "left") +
  tm_compass(type = "8star", position = c("RIGHT", "top")) + 
  tm_layout(legend.frame = TRUE, legend.position = c(0.6, "top"))
sm3
utah = filter(us_states, NAME == "Utah")
zion_bbox = st_as_sfc(st_bbox(nlcd))
sm3_plus = sm3 +
  tm_layout(frame.lwd = 4)

im = tm_shape(utah) +
  tm_polygons(lwd = 3, border.col = "black") + 
  tm_shape(zion_bbox) +
  tm_polygons(col = "green", lwd = 1) +
  tm_layout(title = "UTAH", title.size = 2, title.position = c("center", "center")) +
  tm_layout(frame = FALSE, bg.color = NA)

library(grid)
print(sm3_plus, vp = grid::viewport(0.5, 0.5, width = 0.95, height = 0.95))
print(im, vp = grid::viewport(0.2, 0.4, width = 0.35, height = 0.35))
```
1. Create facet maps of countries in Eastern Africa:
    - with one facet showing HDI and the other representing population growth (hint: using variables `HDI` and `pop_growth` respectively)
    - with a 'small multiple' per country <!-- animated map, interactive map -->
```{r}
eastern_africa = filter(africa, subregion == "Eastern Africa")
tm_shape(eastern_africa) +
  tm_polygons(col = c("HDI", "pop_growth")) +
  qtm(africa, fill = NULL)
# https://github.com/mtennekes/tmap/issues/190#issuecomment-384184801
tm_shape(eastern_africa) +
  tm_polygons("pop_growth", style = "jenks", palette = "BuPu") +
  tm_facets(by = "name", showNA = FALSE)
```
1. Building on the previous facet map examples, create animated maps of East Africa:
    - showing first the spatial distribution of HDI scores then population growth
    - showing each country in order
```{r}
m = tm_shape(eastern_africa) +
  tm_polygons(col = c("HDI", "pop_growth")) +
  qtm(africa, fill = NULL) +
  tm_facets(ncol = 1, nrow = 1)
tmap_animation(m, filename = "m.gif")
browseURL("m.gif")
m = tm_shape(eastern_africa) +
  tm_polygons("pop_growth", style = "jenks", palette = "BuPu") +
  tm_facets(by = "name", showNA = FALSE, ncol = 1, nrow = 1)
tmap_animation(m, filename = "m.gif")
browseURL("m.gif")
```
1. Create an interactive map of Africa:
    - with **tmap**
    - with **mapview**
    - with **leaflet**
    - bonus: for each approach add a legend (if not automatically provided) and a scale bar
1. Sketch on paper ideas for a web mapping app that could be used to make transport or land-use policies more evidence based:
    - In the city you live in, for a couple of users per day
    - In the country you live in, for dozens of users per day
    - Worldwide for hundreds of users per day and large data serving requirements
1. How would app design, deployment and project management decisions change as the scale of map deployment increases?
1. Update the code in `coffeeApp/app.R` so that instead of centering on Brazil the user can select which country to focus on:
    - Using `textInput()`
    - Using `selectInput()`
1. Reproduce Figure \@ref(fig:tmshape) and the 1st and 6th panel of Figure \@ref(fig:break-styles) as closely as possible using the **ggplot2** package.
```{r}
# fig 9.1 (\@ref(fig:tmshape))
ggplot() + 
  geom_sf(data = nz, color = NA) +
  coord_sf(crs = st_crs(nz), datum = NA) + 
  theme_void()
ggplot() +
  geom_sf(data = nz, fill = NA) +
  coord_sf(crs = st_crs(nz), datum = NA) +
  theme_void()
ggplot() +
  geom_sf(data = nz) + 
  coord_sf(crs = st_crs(nz), datum = NA) +
  theme_void()
# fig 9.7 ( \@ref(fig:break-styles))
ggplot() +
  geom_sf(data = nz, aes(fill = Median_income)) +
  coord_sf(crs = st_crs(nz), datum = NA) + 
  scale_fill_distiller(palette = "Oranges", direction = 1) + 
  theme_void()
ggplot() + 
  geom_sf(data = nz, aes(fill = Island)) + 
  coord_sf(crs = st_crs(nz), datum = NA) + 
  scale_fill_brewer(palette = "Set3") + 
  theme_void()
# fig 9.13 (\@ref(fig:urban-facet))
# ggplot() + geom_sf(data = world) + geom_sf(data = urb_1970_2030, aes(size = population_millions)) + coord_sf(crs = st_crs(urb_1970_2030), datum = NA) + facet_wrap(~year, nrow = 2) + theme(panel.background = element_rect(fill = "white"))
```
1. Join `us_states` and `us_states_df` together and calculate a poverty rate for each state using the new dataset.
Next, construct a continuous area cartogram based on total population. 
Finally, create and compare two maps of the poverty rate: (1) a standard choropleth map and (2) a map using the created cartogram boundaries.
What is the information provided by the first and the second map?
How do they differ from each other?
```{r}
library(cartogram)
# prepare the data
us = st_transform(us_states, 2163)
us = us %>% 
  left_join(us_states_df, by = c("NAME" = "state"))
# calculate a poverty rate
us$poverty_rate = us$poverty_level_15 / us$total_pop_15
# create a regular map
ecm1 = tm_shape(us) + tm_polygons("poverty_rate", title = "Poverty rate")
# create a cartogram
us_carto = cartogram(us, "total_pop_15")
ecm2 = tm_shape(us_carto) + tm_polygons("poverty_rate", title = "Poverty rate")
# combine two maps
tmap_arrange(ecm1, ecm2)
```
1. Visualize population growth in Africa. 
Next, compare it with the maps of a hexagonal and regular grid created using the **geogrid** package.
```{r}
# devtools::install_github("jbaileyh/geogrid")
library(geogrid)

hex_cells = calculate_grid(africa, grid_type = "hexagonal", seed = 25, learning_rate = 0.03)
africa_hex = st_as_sf(assign_polygons(africa, hex_cells))

reg_cells = calculate_grid(africa, grid_type = "regular", seed = 25, learning_rate = 0.03)
africa_reg = st_as_sf(assign_polygons(africa, reg_cells))

tgg1 = tm_shape(africa) + tm_polygons("pop_growth", title = "Population's growth (annual %)")
tgg2 = tm_shape(africa_hex) + tm_polygons("pop_growth", title = "Population's growth (annual %)")
tgg3 = tm_shape(africa_reg) + tm_polygons("pop_growth", title = "Population's growth (annual %)")

tmap_arrange(tgg1, tgg2, tgg3)
```

