---
title: "Chapter 12"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
library(spData)
```

# Chapter 12


1. Compute the following terrain attributes from the `dem` datasets loaded with `data("landslides", package = "RSAGA")` with the help of R-GIS bridges (see Chapter \@ref(gis)):
    - slope
    - plan curvature
    - profile curvature
    - catchment area
1. Extract the values from the corresponding output rasters to the `landslides` data frame (`data(landslides, package = "RSAGA"`) by adding new variables called `slope`, `cplan`, `cprof`, `elev` and `log_carea`. Keep all landslide initiation points and 175 randomly selected non-landslide points (see section \@ref(case-landslide) for details).
1. Use the derived terrain attribute rasters in combination with a GLM to make a spatial prediction map similar to that shown in Figure \@ref(fig:lsl-susc).
Running `data("study_mask", package = "spDataLarge")` attaches a mask of the study area.
1. Compute a 100-repeated 5-fold non-spatial cross-validation and spatial CV based on the GLM learner and compare the AUROC values from both resampling strategies with the help of boxplots (see Figure \@ref(fig:boxplot-cv)).
Hint: You need to specify a non-spatial task and a non-spatial resampling strategy.
<!-- @Patrick: talk in person; but I think this step is not necessary since spatial and non-spatial partitions must be different -->
<!-- Before running the spatial cross-validation for both tasks set a seed to make sure that both use the same partitions which in turn guarantees comparability.-->
1. Model landslide susceptibility using a quadratic discriminant analysis [QDA, @james_introduction_2013].
Assess the predictive performance (AUROC) of the QDA. 
What is the difference between the spatially cross-validated mean AUROC value of the QDA and the GLM?
<!-- so I think, setting a seed here makes sure that the same spatial partitions are used for both models, right?-->
Hint: Before running the spatial cross-validation for both learners set a seed to make sure that both use the same spatial partitions which in turn guarantees comparability.
1. Run the SVM without tuning the hyperparameters.
Use the `rbfdot` kernel with $\sigma$ = 1 and *C* = 1. 
Leaving the hyperparameters unspecified in **kernlab**'s `ksvm()` would otherwise initialize an automatic non-spatial hyperparameter tuning.
