---
title: "Chapter 14: Ecology"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions14}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(raster)
library(vegan)
library(mlr)
library(parallelMap)
```

# Chapter 14

1) Run a NMDS using the percentage data of the community matrix. 
Report the stress value and compare it to the stress value as retrieved from the NMDS using presence-absence data.
What might explain the observed difference?

```{r, cache=TRUE}
data("study_area", "random_points", "comm", "dem", "ndvi", package = "RQGIS")
pa = decostand(comm, "pa")
pa = pa[rowSums(pa) != 0, ]
comm = comm[rowSums(comm) != 0, ]

set.seed(25072018)
nmds_pa = metaMDS(comm = pa, k = 4, try = 500)
nmds_per = metaMDS(comm = comm, k = 4, try = 500)
nmds_pa$stress
nmds_per$stress
```

The NMDS using the presence-absence values yields a better result (`nmds_pa$stress`) than the one using percentage data (`nmds_per$stress`).
This might seem surprising at first sight.
On the other hand, the percentag matrix contains both more information and more noise.
Another aspect is how the data was collected.
Imagine a botanist in the field.
It might seem feasible to differentiate between a plant which has a cover of 5% and another species that covers 10%.
However, what about a herbal species that was only detected three times and consequently has a very tiny cover, e.g., 0.0001%. 
Maybe another herbal species was detected 6 times, is its cover then 0.0002%?
The point here is that percentage data as specified during a field campaign might reflect a precision that the data does not have.
This again introduces noise which in turn will worsen the ordination result.
Still, it is a valuable information if one species had a higher frequency or coverage in one plot than another compared to just presence-absence data.
One compromise would be to use a categorical scale such as the Londo scale.

2) Compute all the predictor rasters we have used in the chapter (catchment slope, catchment area), and put them into a raster stack.
Add `dem` and `ndvi` to the raster stack.
Next, compute profile and tangential curvature as additional predictor rasters and add them to the raster stack (hint: `grass7:r.slope.aspect`).
Finally, construct a response-predictor matrix. 
The scores of the first NMDS axis (which were the result when using the presence-absence community matrix) rotated in accordance with elevation represent the response variable, and should be joined to `random_points` (use an inner join).
To complete the response-predictor matrix, extract the values of the environmental predictor raster stack to `random_points`.

```{r, eval=FALSE}
# first compute the terrain attributes we have also used in the chapter
library(RQGIS)
library(dplyr)
library(raster)
library(vegan)

data("dem", package = "RQGIS")
ep = run_qgis(alg = "saga:sagawetnessindex",
              DEM = dem,
              SLOPE_TYPE = 1, 
              SLOPE = tempfile(fileext = ".sdat"),
              AREA = tempfile(fileext = ".sdat"),
              load_output = TRUE,
              show_output_paths = FALSE)
ep = c(dem, ndvi, ep) %>%
  stack()
names(ep) = c("dem", "ndvi", "carea", "cslope")
ep$carea = log10(ep$carea)
# computing the curvatures
get_usage("grass7:r.slope.aspect")
curvs = run_qgis("grass7:r.slope.aspect",
                 elevation = dem,
                 pcurvature = file.path(tempdir(), "pcurvature.tif"),
                 tcurvature = file.path(tempdir(), "tcurvature.tif"),
                 load_output = TRUE)
# adding curvatures to ep
ep = addLayer(ep, curvs$pcurvature, curvs$tcurvature)
random_points[, names(ep)] = raster::extract(ep, as(random_points, "Spatial"))
elev = dplyr::filter(random_points, id %in% rownames(pa)) %>% 
  dplyr::pull(dem)
# rotating NMDS in accordance with altitude (proxy for humidity)
rotnmds = MDSrotate(nmds_pa, elev)
# extracting the first two axes
sc = scores(rotnmds, choices = 1:2)
rp = data.frame(id = as.numeric(rownames(sc)),
                sc = sc[, 1])
# join the predictors (dem, ndvi and terrain attributes)
rp = inner_join(random_points, rp, by = "id")
```

3) Use the response-predictor matrix of the previous exercise to fit a random forest model. 
Find the optimal hyperparameters and use them for making a prediction map.

```{r, eval=FALSE}
data("dem", "study_area", package = "RQGIS")
# TUNING, MODELING AND PREDICTION
#********************************

# create a task
task = makeRegrTask(data = rp, target = "sc", coordinates = coords)
# define the learner
lrn_rf = makeLearner(cl = "regr.ranger", predict.type = "response")
# spatial partitioning
perf_level = makeResampleDesc("SpCV", iters = 5)
# specifying random search
ctrl = makeTuneControlRandom(maxit = 50L)
# specifying the search space
ps = makeParamSet(
  makeIntegerParam("mtry", lower = 1, upper = ncol(rp) - 1),
  makeNumericParam("sample.fraction", lower = 0.2, upper = 0.9),
  makeIntegerParam("min.node.size", lower = 1, upper = 10)
)
# hyperparamter tuning
set.seed(07092018)
tune = tuneParams(learner = lrn_rf,
                  task = task,
                  resampling = perf_level,
                  par.set = ps,
                  control = ctrl,
                  measures = mlr::rmse)
# saveRDS(tune, "extdata/rf_tune_50it.rds")
# define a learner using the optimal hyperparameter combination
lrn_rf = makeLearner(cl = "regr.ranger",
                     predict.type = "response",
                     mtry = tune$x$mtry,
                     sample.fraction = tune$x$sample.fraction,
                     min.node.size = tune$x$min.node.size)
# doing the same more elegantly using setHyperPars()
# lrn_rf = setHyperPars(makeLearner("regr.ranger", predict.type = "response"),
#                       par.vals = tune$x)
# train model
model_rf = train(lrn_rf, task)

# make the prediction, but before that check if there are NAs in our raster
# stack...
plot(is.na(ep))
# there are, but these just correspond to the outer frame, hence we can delete
# them safely. Otherwise, the predict function will complain later on
ep = trim(ep$pcurvature, values = NaN) %>%
  crop(ep, .)
# check if it worked
is.na(ep)  # no NAs, perfect
# convert raster stack into a dataframe
new_data = as.data.frame(as.matrix(ep))
# apply the model to the dataframe
pred_rf = predict(model_rf, newdata = new_data, na.rm = TRUE)

# put the predicted values into a raster
pred = ep$dem
# replace altitudinal values by rf-prediction values
pred[] = pred_rf$data$response

# PREDICTION MAP
#***************

library("latticeExtra")
library("grid")

# create a color palette
blue = rgb(0, 0, 146, maxColorValue = 255)
lightblue = rgb(0, 129, 255, maxColorValue = 255)
turquoise = rgb(0, 233, 255, maxColorValue = 255)
green = rgb(142, 255, 11, maxColorValue = 255)
yellow = rgb(245, 255, 8, maxColorValue = 255)
orange = rgb(255, 173, 0, maxColorValue = 255)
lightred = rgb(255, 67, 0, maxColorValue = 255)
red = rgb(170, 0, 0, maxColorValue = 255)
pal = colorRampPalette(c(blue, lightblue, turquoise, green, yellow,
                         orange, lightred, red))

# restrict the prediction to the study area
pred = mask(pred, study_area) %>%
  trim

# create a hillshade
hs = hillShade(terrain(dem), terrain(dem, "aspect")) %>%
  mask(., study_area)
spplot(extend(pred, 2), col.regions = pal(50), alpha.regions = 0.7,
       scales = list(draw = TRUE,
                     tck = c(1, 0),
                     cex = 0.8),
       colorkey = list(space = "right", width = 0.5, height = 0.5,
                       axis.line = list(col = "black")),
       sp.layout = list(
         # list("sp.points", as(random_points, "Spatial"), pch = 16,
         #      col = "black", cex = 0.8, first = FALSE),
         list("sp.polygons", as(study_area, "Spatial"),
              col = "black", first = FALSE, lwd = 3)
       )
) +
  latticeExtra::as.layer(spplot(hs, col.regions = gray(0:100 / 100)),
                         under = TRUE)
```

4) Retrieve the bias-reduced RMSE of a random forest model using spatial cross-validation including the estimation of optimal hyperparameter combinations (random search with 50 iterations) in an inner tuning loop (see [section](https://geocompr.robinlovelace.net/spatial-cv.html#svm)).
Parallelize the tuning level (see section [section](https://geocompr.robinlovelace.net/spatial-cv.html#svm)).
Report the mean RMSE and use a boxplot to visualize all retrieved RMSEs.

```{r, eval=FALSE}
# extract the coordinates into a separate dataframe
coords = sf::st_coordinates(rp) %>%
  as.data.frame %>%
  rename(x = X, y = Y)
# only keep response and predictors which should be used for the modeling
rp = dplyr::select(rp, -id, -spri) %>%
  st_set_geometry(NULL)

# create task
task = makeRegrTask(data = rp, target = "sc", coordinates = coords)
# learner
lrn_rf = makeLearner(cl = "regr.ranger", predict.type = "response")
# performance estimation level with 5 spatial partitions and 100 repetitions
perf_level = makeResampleDesc(method = "SpRepCV", folds = 5, reps = 100)
# five spatially disjoint partitions in the tune level (one repetition)
tune_level = makeResampleDesc(method = "SpCV", iters = 5)
# random search with 50 iterations
ctrl = makeTuneControlRandom(maxit = 50)
# specifying the search space
ps = makeParamSet(
  makeIntegerParam("mtry", lower = 1, upper = ncol(rp) - 1),
  makeNumericParam("sample.fraction", lower = 0.2, upper = 0.9),
  makeIntegerParam("min.node.size", lower = 1, upper = 10)
)
# wrap it all up
wrapped_lrn_rf = makeTuneWrapper(learner = lrn_rf,
                                   # inner loop (tunning level)
                                   resampling = tune_level,
                                   # hyperparameter seach space
                                   par.set = ps,
                                   # random search
                                   control = ctrl,
                                   show.info = TRUE,
                                   # performance measure
                                   measures = mlr::rmse)
# make sure that the modeling goes on even if one model fails
configureMlr(on.learner.error = "warn", on.error.dump = TRUE)

# initialize parallelization
if (Sys.info()["sysname"] %in% c("Linux", "Darwin")) {
  parallelStart(mode = "multicore",
                # parallelize the hyperparameter tuning level
                level = "mlr.tuneParams",
                # just use half of the available cores
                cpus = round(parallel::detectCores() / 2),
                mc.set.seed = TRUE)
}

if (Sys.info()["sysname"] == "Windows") {
  parallelStartSocket(level = "mlr.tuneParams",
                      cpus =  round(parallel::detectCores() / 2))
}

# run the spatial cross-validation
set.seed(12345)
result_rf = mlr::resample(learner = wrapped_lrn_rf,
                          task = task,
                          resampling = perf_level,
                          extract = getTuneResult,
                          measures = mlr::rmse)
# stop parallelization
parallelStop()
# save your result, e.g.:
# saveRDS(result_rf, file = "extdata/rf_sp_sp_50it.rds")

# Visualization of non-spatial overfitting
boxplot(result_rf$measures.test$rmse, ylab = "RMSE")
```

5) Retrieve the bias-reduced RMSE of a simple linear model using spatial cross-validation. 
Compare the result to the result of the random forest model by making RMSE boxplots for each modeling approach.

```{r, eval=FALSE}
# create a task
task = makeRegrTask(data = rp, target = "sc",
                    coordinates = coords)
# run listLearners to find out which models could be thrown at our task
# lrns = listLearners(task, warn.missing.packages = FALSE)
# dplyr::select(lrns, class, name, short.name, package)

# define a learner
lrn = makeLearner(cl = "regr.lm", predict.type = "response")
# simple lm of the stats package
# getLearnerPackages(lrn)
# helpLearner(lrn)
# so the model being fitted is simply a lm
# getLearnerModel(train(lrn, task))

# performance level
perf_level = makeResampleDesc(method = "SpRepCV", folds = 5, reps = 100)
result_lm = mlr::resample(learner = lrn,
                          task = task,
                          resampling = perf_level,
                          measures = mlr::rmse)
# save your result, e.g.:
# saveRDS(result_lm, "extdata/lm_sp.rds")
boxplot(result_rf$measures.test$rmse, result_lm$measures.test$rmse,
        col = c("lightblue2", "mistyrose2"),
        names = c("random forest", "lm"), ylab = "RMSE")
```

In fact, `lm` performs better than a random forest model.
But keep in mind that the used dataset is small in terms of observations and predictors and that the response-predictor relationships are also relatively linear.
