---
title: "Chapter 7: Geographic data I/O"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(raster)
library(tidyverse)
library(spData)
```

# Chapter 7

1. List and describe three types of vector, raster, and geodatabase formats.

Vector formats: Shapefile (old format supported by many programs), GeoPackage (more recent format with better support of attribute data) and GeoJSON (common format for web mapping).

Raster formats: GeoTiff, Arc ASCII, R-raster (see book for descriptions).

Database formats: PostGIS, SQLite, FileGDB (see book for details).

2. Name at least two differences between `read_sf()` and the more well-known function `st_read()`.

`st_read()` prints outputs and keeps strings as text strings (`st_read()` creates factors). This can be seen from the source code of `read_sf()`, which show's it wraps `st_read()`:

```{r}
read_sf
```


3. Read the `cycle_hire_xy.csv` file from the **spData** package as a spatial object (Hint: it is located in the `misc\` folder).
What is a geometry type of the loaded object? 

```{r}
c_h = read.csv(system.file("misc/cycle_hire_xy.csv", package = "spData")) %>% 
  st_as_sf(coords = c("X", "Y"))
c_h
```

The loaded object has the POINT geometry.

4. Download the borders of Germany using **rnaturalearth**, and create a new object called `germany_borders`.
Write this new object to a file of the GeoPackage format.

```{r, eval=FALSE}
library(rnaturalearth)
germany_borders = ne_countries(country = "Germany", returnclass = "sf")
plot(germany_borders)
st_write(germany_borders, "germany_borders.gpkg")
```

5. Download the global monthly minimum temperature with a spatial resolution of five minutes using the **raster** package.
Extract the June values, and save them to a file named `tmin_june.tif` file (hint: use `raster::subset()`).

```{r, eval=FALSE}
gmmt = getData("worldclim", var = "tmin", res = 5)
gmmt_june = raster::subset(gmmt, "tmin6")
writeRaster(gmmt_june, "tmin_june.tif")
```

6. Create a static map of Germany's borders, and save it to a PNG file.

```{r, eval=FALSE}
png(filename = "germany.png", width = 350, height = 500)
plot(st_geometry(germany_borders), axes = TRUE, graticule = TRUE)
dev.off()
```

7. Create an interactive map using data from the `cycle_hire_xy.csv` file. 
Export this map to a file called `cycle_hire.html`.

```{r, eval=FALSE}
library(mapview)
mapview_obj = mapview(c_h, zcol = "nbikes", legend = TRUE)
mapshot(mapview_obj, file = "cycle_hire.html")
```

