---
title: "Chapter 4"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geocompr-solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Prerequisites {-}

The solutions assume the following packages are attached (other packages will be attached when needed):

```{r packages, message=FALSE, warning=FALSE}
library(sf)
library(raster)
library(tidyverse)
library(spData)
library(spDataLarge)
library(RStoolbox)
```

# Chapter 4

<!-- vector exercises -->
1. It was established in section \@ref(spatial-vec) that Canterbury was the region of New Zealand containing most of 100 highest points in the country. How many of these high points does Canterbury Region contain?
```{r}
library(tmap)
tmap_mode("view")
qtm(nz) + qtm(nz_height)
canterbury = nz %>% filter(Name == "Canterbury Region")
canterbury_height = nz_height[canterbury, ]
nrow(canterbury_height) # answer: 61
```
1. Which region has the second highest number of `nz_height` points in, and how many does it have?
```{r}
nz_height_count = aggregate(nz_height, nz, length)
nz_height_combined = cbind(nz, count = nz_height_count$elevation)
nz_height_combined %>% 
  st_set_geometry(NULL) %>% 
  dplyr::select(Name, count) %>% 
  arrange(desc(count)) %>% 
  na.omit()
```
1. Generalizing the question to all regions: how many of New Zealand's 16 regions contain points which belong to the top 100 highest points in the country? Which regions?
    - Bonus: create a table listing these regions in order of the number of points and their name.
<!-- Raster exercises-->
4. Use `data(dem, package = "RQGIS")`, and reclassify the elevation in three classes: low, middle and high.
Secondly, compute the NDVI (`data(ndvi, package = "RQGIS")`) and the mean elevation for each altitudinal class.
```{r}
library(classInt)
data(dem, package = "RQGIS")
data(ndvi, package = "RQGIS")
summary(dem)
# find quantile breaks
brk = classIntervals(values(dem), n = 3)$brk
# also try
# breask = classIntervals(values(dem), n = 3, style = "fisher")
# construct reclassification matrix
rcl = matrix(c(brk[1] - 1, brk[2], 1, brk[2], brk[3], 2, brk[3], brk[4], 3), 
             ncol = 3, byrow = TRUE)
# reclassify altitudinal raster
recl = reclassify(dem, rcl = rcl)
# compute the mean dem and ndvi values for each class
zonal(stack(dem, ndvi), recl, fun = "mean")
```

5. Apply a line detection filter to `data(dem, package = "RQGIS")`.
```{r}
# from the focal help page:
# Laplacian filter: filter=matrix(c(0,1,0,1,-4,1,0,1,0), nrow=3)
# Sobel filter: filter=matrix(c(1,2,1,0,0,0,-1,-2,-1) / 4, nrow=3)
# compute the Sobel filter
# check if there are NAs
is.na(dem)  # just 0s, so no NAs
sobel = focal(dem, w = matrix(c(1, 2, 1, 0, 0, 0, -1, -2, -1) / 4, nrow = 3))
# CHECK IF THIS IS CORRECT
```

6. Calculate the NDVI of a Landsat image. 
Use the Landsat image provided by the **spDataLarge** package (`system.file("raster/landsat.tif", package="spDataLarge")`).

```{r}
file = system.file("raster/landsat.tif", package = "spDataLarge")
r = stack(file)
# compute NDVI manually
ndvi = (r[["landsat.4"]] - r[["landsat.3"]]) / (r[["landsat.4"]] + r[["landsat.3"]])
# compute NDVI with the help of RStoolbox
ndvi_rstoolbox = spectralIndices(r, red = 3, nir = 4, indices = "NDVI")
all.equal(ndvi, ndvi_rstoolbox)
```

7. This [post](https://stackoverflow.com/questions/35555709/global-raster-of-geographic-distances) shows how to compute distances to the nearest coastline using `raster::distance()`.
Retrieve a digital elevation model of Spain, and compute a raster which represents the distance to the coast.
(Hint: Have a look at `getData()` to retrieve a digital elevation model and administrative boundaries for Spain.)
Before computing the distance raster, you might want to increase the resolution of the input dem raster, otherwise computing time might become too long. 
Secondly, weight the distance raster with elevation.
Every 100 altitudinal meters should increase the distance to the coast by 10 km.
Finally, compute the difference between the raster using the euclidean distance and the raster weighted by elevation.
(Note that this is a very simple weighting approach.
A more advanced approach might instead weight by flow direction, i.e. favor the steepest drop or the slightest increase in elevation.)

```{r}
library(raster)
# find out the ISO_3 code of Spain
dplyr::filter(ccodes(), NAME %in% "Spain")
# retrieve a dem of Spain
dem = getData("alt", country = "ESP", mask = FALSE)
# change the resolution to decrease computing time
agg = aggregate(dem, fact = 5)
poly = getData("GADM", country = "ESP", level = 1)
plot(dem)
plot(poly, add = TRUE)
# visualize NAs
plot(is.na(agg))
# construct a distance input raster
# we have to set the land cells to NA and the sea cells to an arbitrary value since 
# raster::distance computes the distance to the nearest non-NA cell
dist = is.na(agg)
cellStats(dist, summary)
# convert land cells into NAs and sea cells into 1s
dist[dist == FALSE] = NA
dist[dist == TRUE] = 1
plot(dist)
# compute distance to nearest non-NA cell
dist = raster::distance(dist)
# just keep Spain
dist = mask(dist, poly)
# convert distance into km
dist = dist / 1000
# now let's weight each 100 altitudinal meters by an additionaly distance of 10 km
agg = mask(agg, poly)
agg[agg < 0] = 0
weight = dist + agg / 100 * 10
plot(weight - dist)
```
