---
title: "Map styles available in the tmap package"
author: "Robin Lovelace, Jakub Nowosad, Jannes Muenchow"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
      toc: true
      number_sections: true
vignette: >
  %\VignetteIndexEntry{tmap-styles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE,
                      fig.width = 8, fig.height = 4)
```

This vignette builds on the [making maps chapter](https://geocompr.robinlovelace.net/adv-map.html) of [the Geocomputation with R book](https://geocompr.github.io/).
Its goal is to demonstrate all possible map styles available in the **tmap** package.

## Prerequisites {-}

The examples below assume the following packages are attached:

```{r}
library(spData) # example datasets
library(tmap)   # map creation
library(sf)     # spatial data reprojection
```

The `world` object containing a world map data from [Natural Earth](https://www.naturalearthdata.com/) and information about countries' names, regions and subregions they belong to, areas, life expectancies, and populations.
This object is in geographical coordinates using the WGS84 datum, however for mapping purposes the Mollweide projection is a better alternative (learn more in [the modifying map projections section](https://geocompr.robinlovelace.net/reproj-geo-data.html#modifying-map-projections)).
The `st_tranform` function from the **sf** package allows for quick reprojection to the selected coordinate reference system (e.g., the Mollweide projection can be represented as `"+proj=moll"`).

```{r}
world_moll = st_transform(world, crs = "+proj=moll")
```

# One color

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lightblue")
```

# Coloring of adjacent polygons

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "MAP_COLORS")
```

`tmaptools::palette_explorer()`

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "MAP_COLORS",
              palette = "Pastel1")
```

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "MAP_COLORS",
              minimize = TRUE)
```

https://en.wikipedia.org/wiki/Four_color_theorem

# Categorical maps

`tm_polygons(col = "subregion", style = "cat")`

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "subregion")+
  tm_layout(legend.outside = TRUE) 
```

# Discrete maps

## Fixed

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp", 
              style = "fixed",
              breaks = c(50, 60, 70, 80, 90),
              legend.hist = TRUE) +
  tm_layout(legend.outside = TRUE) 
```

## Pretty

The `"pretty"` style rounds breaks into whole numbers where possible and spaces them evenly.

`tm_polygons(col = "lifeExp", style = "pretty")`

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp",
              legend.hist = TRUE) +
  tm_layout(legend.outside = TRUE) 
```

<!-- n is approximation -->

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp",
              legend.hist = TRUE,
              n = 4) +
  tm_layout(legend.outside = TRUE) 
```

<!-- https://r-spatial.github.io/classInt/index.html -->


## Breaks based on the centred and scaled variable

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp", 
              style = "sd",
              legend.hist = TRUE) +
  tm_layout(legend.outside = TRUE)
```


## Jenks natural breaks 

The `"jenks"` style identifies groups of similar values in the data and maximizes the differences between categories.

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp",
              style = "jenks",
              legend.hist = TRUE) +
  tm_layout(legend.outside = TRUE) 
```

## Fisher-Jenks algorithm

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp",
              style = "fisher",
              legend.hist = TRUE) +
  tm_layout(legend.outside = TRUE) 
```

## Bagged clustering

```{r, message=FALSE}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp",
              style = "bclust",
              legend.hist = TRUE) +
  tm_layout(legend.outside = TRUE) 
```

## Hierarchical clustering

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp",
              style = "hclust",
              legend.hist = TRUE) +
  tm_layout(legend.outside = TRUE) 
```

## k-means clustering

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp", 
              style = "kmeans",
              legend.hist = TRUE) +
  tm_layout(legend.outside = TRUE) 
```

## Quantile breaks

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp", 
              style = "quantile",
              legend.hist = TRUE) +
  tm_layout(legend.outside = TRUE) 
```

## Equal breaks

The `"equal"` style divides input values into bins of equal range, and is appropriate for variables with a uniform distribution (not recommended for variables with a skewed distribution as the resulting map may end-up having little color diversity).

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp", 
              style = "equal",
              legend.hist = TRUE) +
  tm_layout(legend.outside = TRUE) 
```


## Continuous

The `"cont"` style presents a large number of colors over continuous color field.

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp",
              style = "cont") +
  tm_layout(legend.outside = TRUE) 
```

The `"order"` style also presents a large number of colors over continuous color field. 
However, this style is suited to visualize skewed distributions.

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "lifeExp",
              style = "order") +
  tm_layout(legend.outside = TRUE) 
```

## Logarithmic scale

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "pop") +
  tm_layout(legend.outside = TRUE) 
```

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "pop", 
              style = "cont") +
  tm_layout(legend.outside = TRUE) 
```

The `"log10_pretty"` style 

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "pop", 
              style = "log10_pretty") +
  tm_layout(legend.outside = TRUE) 
```

The `"log10"` style 

```{r}
tm_shape(world_moll) +
  tm_polygons(col = "pop", 
              style = "log10") +
  tm_layout(legend.outside = TRUE) 
```
