---
output: github_document
# output: word_document
bibliography: ../vignettes/refs.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, eval=FALSE, echo=FALSE}
# generate/publish online the materials
knitr::purl("README.Rmd")

# save html slides
wd_old = setwd("slides")
f = list.files(pattern = "spatial")
zip(zipfile = "spatial.zip", files = f)
f = list.files(pattern = "roadworks")
zip(zipfile = "roadworks.zip", files = f)
setwd(wd_old)

mv -v slides/*.zip . # in bash
```

## Aim

The aim of Geocomputation and Data Analysis with R is to get you up-to-speed with high performance geographic processing, analysis, visualisation and modelling capabilities from the command-line.
The course will be delivered in R, a statistical programming language popular in academia, industry and, increasing, the public sector.
It will teach a range of techniques using recently developments in the package [**sf**](https://github.com/r-spatial/sf) and the 'metapackage' [**tidyverse**](https://www.tidyverse.org/), based on the open source book [Geocomputation with R](https://geocompr.robinlovelace.net/) [@lovelace_geocomputation_2019]. 

## Location

Leeds Institute for Data Analytics, 40 University Road, LS2 9JT.
It can be seen on [OpenStreetMap](https://www.openstreetmap.org/way/84749920), on the [LIDA website](https://lida.leeds.ac.uk/about-lida/contact/) and highlighted in red below (it's a 20 minute walk from the station):

```{r itsmap, include=FALSE}
library(osmdata)
library(tmap)
lida = opq("Leeds") %>% 
  add_osm_feature(key = "name", value = "Worsley Building") %>% 
  osmdata_sf()
# m = mapview::mapview(lida$osm_polygons, zoom = 16)
lida_centroid = lida$osm_polygons %>% 
  sf::st_centroid()
lon = sf::st_coordinates(lida_centroid)[1]
lat = sf::st_coordinates(lida_centroid)[2]
m = tm_shape(lida$osm_polygons) +
  tm_polygons(col = "red") +
  tm_view(set.view = c(lon + 0.005, lat - 0.005, 15)) +
  tm_basemap(server = leaflet::providers$CartoDB.Positron)
tmap_mode("view")
ml = tmap_leaflet(m)
```

```{r, echo=FALSE}
ml
```

## Course contents

### Day 1: Foundations

09:00-09:30 Arrival and set-up

09:30-11:00 Introduction to the course and software

- Introduction to R
- R installation questions/debugging
- How to use RStudio (practical in groups of 2)
- R classes and working with data frames (CC)

11:15-12:30: TBC

<!-- And example from the PCT -->

**Lunch**

13:30-15:00 Spatial data in R

- Spatial data in R
- R's spatial ecosystem (see section [1.4 of Geocomputation with R - package ecosystem](https://geocompr.robinlovelace.net/intro.html#rs-spatial-ecosystem))
- Practical: [Section 3.2 to 3.2.2](https://geocompr.robinlovelace.net/attr.html#vector-attribute-manipulation) of handouts
- [Exercises](https://geocompr.robinlovelace.net/attr.html#exercises-1): 1 to 3 in the hand-outs
- Further reading: [Chapter 7](https://geocompr.robinlovelace.net/read-write.html) of Geocomputation with R - data import/export and [exercises](https://geocompr.robinlovelace.net/read-write.html)
- Bonus (relies on internet/data): read-in and analyse datasets you use at work as data frames or sf objects. 
  
15:15-16:30 TBC

### Day 2 GIS, visualisation, modelling

09:30-11:00 TBC

11:15-12:30: TBC

<!-- And example from the PCT -->

**Lunch**

13:30-15:00 TBC

15:15-16:30 TBC

## Prerequisites

### Prior reading/experience

If you are new to R, ensure you have attempted a basic introductory course such as DataCamp's [introduction to R](https://www.datacamp.com/courses/free-introduction-to-r) free course or equivalent.

If you're interested in R for 'data science' and installing/updating/choosing R packages, these additional resources are recommended (these optional resources are all freely availble online):

- The introductory chapter of [R for Data Science](https://r4ds.had.co.nz/introduction.html)
- Chapter 2 on [setting-up R](https://csgillespie.github.io/efficientR/set-up.html) and section 4.4 on [package selection](https://csgillespie.github.io/efficientR/workflow.html#package-selection) in the book *Efficient R Programming*
### Computing requirements

Attendees are expected to bring their own laptop with the following packages installed and working.
You can check these are all installed, and install those that are not installed, as follows (you can also just type `install.packages("sf")` etc):

```{r packages, message=FALSE, warning=FALSE, eval=FALSE}
install.packages("osmdata")   # for working with open street map data
install.packages("sf")        # a package for working with spatial data
install.packages("spData")    # provides example data
install.packages("stplanr")   # a transport data package
install.packages("tidyverse") # metapackage for data science
install.packages("tmap")      # a mapping package

# Make sure your packages are up-to-date with:
update.packages()
```

## Reproducible example

The code in the following example checks you have the necessary packages installed.
It results in a map that will guide you to the location of the course.

Attach the packages:

```{r, message=FALSE}
library(sf)
library(stplanr)
library(tidyverse)
```

The overall route assuming you're travelling from London:

```{r}
uk = spData::world %>% 
  filter(name_long == "United Kingdom")
origin_lnd = c(-0.1, 51.5)
destination = c(-1.55, 53.8)
odmatrix = matrix(c(origin_lnd, destination), ncol = 2, byrow = TRUE)
line_lnd = st_linestring(odmatrix) %>% 
  st_sfc() %>% 
  st_sf(crs = 4326)
plot(st_geometry(uk), col = "grey")
plot(line_lnd, add = TRUE, lwd = 5, col = "red")
```

## References
