---
output: github_document
# output: word_document
bibliography: ../vignettes/refs.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, eval=FALSE, echo=FALSE}
# generate/publish online the materials
knitr::purl("README.Rmd")

# save html slides
wd_old = setwd("slides")
f = list.files(pattern = "spatial")
zip(zipfile = "spatial.zip", files = f)
f = list.files(pattern = "roadworks")
zip(zipfile = "roadworks.zip", files = f)
setwd(wd_old)

mv -v slides/*.zip . # in bash
```

## Geocomputation and Data Analysis with R

Geocomputation and Data Analysis with R is a 2 day course to be held on the 25^th^ and 26^th^ April 2019 in the Leeds Institute for Data Analytics (LIDA).

## Aim

The aim of the course is to get you up-to-speed with high performance geographic processing, analysis and visualisation capabilities from the command-line.
The course will be delivered in R, a statistical programming language popular in academia, industry and, increasing, the public sector.
It will teach a range of techniques using recently developments in the package [**sf**](https://github.com/r-spatial/sf) and the 'metapackage' [**tidyverse**](https://www.tidyverse.org/), based on the open source book [Geocomputation with R](https://geocompr.robinlovelace.net/) [@lovelace_geocomputation_2019]. 

## Learning objectives

By then end of the course participants should:

- Be able to use R and RStudio as a powerful Geographic Information System (GIS)
<!-- What the different panels within RStudio are and how to use the view panel interactively -->
- Know how R's spatial capabilities fit within the landscape of open source GIS software
<!-- What are the alternatives to R? -->
<!-- What are it's strengths/weaknesses? -->
<!-- How can R interface to dedicated GIS software like QGIS? -->
- Be confident with using R's command-line interface (CLI) and scripting capabilities for geographic data processing
<!-- autocompletion -->
<!-- using help -->
<!-- knowing where to search and create a reproducible example to ask questions on a community forum -->
<!-- source vs console editor -->
<!-- projects and scripts -->
<!-- functions / algorithms -->
- Understand how to import a range of data sources into R
<!-- read-in a shapefile -->
<!-- write-out a geojson -->
<!-- 7.2 Retrieving open data -->
<!-- 7.3 Geographic data packages -->
<!-- 7.4 Geographic web services -->
<!-- 7.5 File formats -->
<!-- 7.6 Data Input (I) -->
<!-- 7.7 Data output (O) -->
- Be able to perform a range of attribute operations such as subsetting and joining
<!-- 3.2.1 Vector attribute subsetting -->
<!-- 3.2.2 Vector attribute aggregation -->
<!-- 3.2.3 Vector attribute joining -->
<!-- 3.2.4 Creating attributes and removing spatial information -->
- Understand how to implement a range of spatial data operations including spatial subsetting and spatial aggregation
<!-- 4.2.1 Spatial subsetting -->
<!-- 4.2.2 Topological relations -->
<!-- 4.2.3 Spatial joining -->
<!-- 4.2.4 Non-overlapping joins -->
<!-- 4.2.5 Spatial data aggregation -->
- Have the confidence to output the results of geographic research in the form of static and interactive maps
<!--     8.2.1 tmap basics -->
<!--     8.2.2 Map objects -->
<!--     8.2.3 Aesthetics -->
<!--     8.2.4 Color settings -->
<!--     8.2.5 Layouts -->
<!--     8.2.6 Faceted maps -->
<!-- 8.4 Interactive maps -->

## Location

Leeds Institute for Data Analytics, 40 University Road, LS2 9JT.
It can be seen on [OpenStreetMap](https://www.openstreetmap.org/way/84749920), on the [LIDA website](https://lida.leeds.ac.uk/about-lida/contact/) and highlighted in red below (it's a 20 minute walk from the station):

```{r itsmap, include=FALSE}
library(leaflet)
library(osmdata)
library(tmap)
lida = opq("Leeds") %>% 
  add_osm_feature(key = "name", value = "Worsley Building") %>% 
  osmdata_sf()
# m = mapview::mapview(lida$osm_polygons, zoom = 16)
lida_centroid = lida$osm_polygons %>% 
  sf::st_centroid()
lon = sf::st_coordinates(lida_centroid)[1]
lat = sf::st_coordinates(lida_centroid)[2]
m = tm_shape(lida$osm_polygons) +
  tm_polygons(col = "red") +
  tm_view(set.view = c(lon + 0.005, lat - 0.005, 15)) +
  tm_basemap(server = leaflet::providers$CartoDB.Positron) +
  tm_minimap(server = leaflet::providers$OpenTopoMap, zoomLevelOffset = -10) +
  tm_scale_bar()
tmap_mode("view")
ml = tmap_leaflet(m)
```

```{r, echo=FALSE}
ml
```

## Course contents

### Day 1: Foundations

09:00-09:30 Arrival and set-up

09:30-11:00 Introduction to the course and software

- Introduction to R
- R installation questions/debugging
- How to use RStudio (practical in groups of 2)
- R classes and working with data frames (CC)

11:15-12:30: TBC

<!-- And example from the PCT -->

**Lunch**

13:30-15:00 Spatial data in R

- Spatial data in R
- R's spatial ecosystem (see section [1.4 of Geocomputation with R - package ecosystem](https://geocompr.robinlovelace.net/intro.html#rs-spatial-ecosystem))
- Practical: [Section 3.2 to 3.2.2](https://geocompr.robinlovelace.net/attr.html#vector-attribute-manipulation) of handouts
- [Exercises](https://geocompr.robinlovelace.net/attr.html#exercises-1): 1 to 3 in the hand-outs
- Further reading: [Chapter 7](https://geocompr.robinlovelace.net/read-write.html) of Geocomputation with R - data import/export and [exercises](https://geocompr.robinlovelace.net/read-write.html)
- Bonus (relies on internet/data): read-in and analyse datasets you use at work as data frames or sf objects. 
  
15:15-16:30 TBC

### Day 2 GIS, visualisation, modelling

09:30-11:00 TBC

11:15-12:30: TBC

<!-- And example from the PCT -->

**Lunch**

13:30-15:00 TBC

15:15-16:30 TBC

## Prerequisites

### Prior experience

**Attendees should already have experience with R.**
In addition to being able to complete introductory course such as DataCamp's [introduction to R](https://www.datacamp.com/courses/free-introduction-to-r) free course or equivalent, attendees should **already be using R** for their work.
This is not an introductory R course but a course for R users wanting to develop their geospatial analysis skills.
If you do not have R experience but already use GIS software and have a strong understanding for geographic data types, and some programming experience, the course may also be appropriate.

In preparation for the course we recommend reading and running the code in these free online resources:

- The introductory chapter of [R for Data Science](https://r4ds.had.co.nz/introduction.html)
- Chapter 2 on [setting-up R](https://csgillespie.github.io/efficientR/set-up.html) and section 4.4 on [package selection](https://csgillespie.github.io/efficientR/workflow.html#package-selection) in the book *Efficient R Programming*

### Computing requirements

**Attendees are expected to bring their own laptop with the following packages installed and working.**
You can check these are all installed, and install those that are not installed, as follows (you can also just type `install.packages("sf")` etc):

```{r packages, message=FALSE, warning=FALSE, eval=FALSE}
install.packages("osmdata")   # for working with open street map data
install.packages("sf")        # a package for working with spatial data
install.packages("spData")    # provides example data
install.packages("stplanr")   # a transport data package
install.packages("tidyverse") # metapackage for data science
install.packages("tmap")      # a mapping package

# Make sure your packages are up-to-date with:
update.packages()
```

## Reproducible example

The code in the following example checks you have the necessary packages installed.
It results in a map that will guide you to the location of the course.

Attach the packages:

```{r, message=FALSE}
library(sf)
library(stplanr)
library(tidyverse)
```

The overall route assuming you're travelling from London:

```{r}
uk = spData::world %>% 
  filter(name_long == "United Kingdom")
origin_lnd = c(-0.1, 51.5)
destination = c(-1.55, 53.8)
odmatrix = matrix(c(origin_lnd, destination), ncol = 2, byrow = TRUE)
line_lnd = st_linestring(odmatrix) %>% 
  st_sfc() %>% 
  st_sf(crs = 4326)
plot(st_geometry(uk), col = "grey")
plot(line_lnd, add = TRUE, lwd = 5, col = "red")
```

## References
